---
title: "COMPASS_TEMPEST_TGW_Collate_AllData"
author: "Stephanie J. Wilson"
date: "2025-09-02"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(ggpubr)
library(dplyr)
library(plyr)
library(tidyverse)
library(tidyr)
library(kableExtra)
library(magick)
library(gt)

```

## Read in all processed file and clean up / add info
```{r, include=FALSE}
#read in all data: 
dat22 <- read.csv("TEMPEST_TGW_2022/COMPASS_TEMPEST_TGW_2022_AllData.csv", colClasses = "character")
dat23 <- read.csv("TEMPEST_TGW_2023/COMPASS_TEMPEST_TGW_2023_AllData.csv", colClasses = "character")
dat24 <- read.csv("TEMPEST_TGW_2024/COMPASS_TEMPEST_TGW_2024_AllData.csv", colClasses = "character")

#combine all the data together into one file 
alldat <- bind_rows(dat22, dat23, dat24)

```
##Add in metadata
```{r, include = FALSE}

metadata <- read.csv("TEMPEST_TGW_Sample_Locations.csv")

metadata <- metadata %>% 
    select(c(Sapflux_ID, Tree_ID, spp, Well_Height_m))


alldat1 <- alldat %>% 
    left_join(metadata, by = c("Sapflux_Tree_ID" = "Sapflux_ID"))

```


## Clean up any columns that need it
```{r, include=FALSE}

##Clean up data
names(alldat1)
final_dat <- alldat1 %>%
    select(-X) %>% 
    dplyr::rename(Sapflux_ID = Sapflux_Tree_ID, 
           Tree_Species = spp) %>% 
    mutate(Plot = str_replace(Plot, "Seawater", "Saltwater")) %>%
    mutate(Time = gsub(":", "", Time)) %>% #fix times so they're all HHMM
    mutate(CH4_Flag = gsub("Within Std Curve Range", "", CH4_Flag)) %>% 
    mutate(CO2_Flag = gsub("Within Std Curve Range", "", CO2_Flag)) %>% 
    mutate(CH4_Flag = gsub("Over Std Curve Range", "adl", CH4_Flag)) %>% 
    mutate(CO2_Flag = gsub("Over Std Curve Range", "adl", CO2_Flag)) %>%     
    mutate(Project = "COMPASS", Site = "TEMPEST") %>% 
    select("Project", "Site", "Year", "Month", "Day", "Plot", "Sample_Grid", "Sapflux_ID", "Tree_ID", "Tree_Species", "Well_Height_m", "Event_Stamp", "Time", "Time_Zone", "Sample_Matrix", "CH4_ppm", "CH4_Flag", "CO2_ppm", "CO2_Flag", "Field_Notes", "Lab_Notes")


```


## Make a metadata file that explains these data
```{r, include=FALSE}

#create a dataframe as the metadata from the columns of alldat 
metadat <- data.frame(Column_Names = colnames(final_dat))

metadat$Description <- ""
metadat$Units <- ""
metadat$Instrument_ifapplicable <- ""

colnames(final_dat)
#add information about each line: 
metadat$Description <- c("Project name", 
                         "Site where sample collected", 
                         "Year of sample collection", 
                         "Month of sample collection", 
                         "Day of sample collection", 
                         "Plot where sample collected (SW = Saltwater, FW = Freshwater, C = Control, Estuary (Sourcewater))", 
                         "Grid where sample collected (Letter + Number)", 
                         "ID of sapflux sensors on tree that sample came from", 
                         "ID of tree (# on metal tag on tree)", 
                         "Tree Species (LITU = Liriodendron tulipifera, ACRU = Acer rubrum, FAGR = Fagus grandifolia)", 
                         "Height of tree gas well from ground", 
                         "Event sampling timestamp (#, AM/PM, D#, Pre/Post)", 
                         "Time of sample collection", 
                         "Time zone of sample collection", 
                         "Matrix that sample was collected from (G = gas, W = water)", 
                         "Methane (CH4) concentration", 
                         "Methane (CH4) concentration flag (bdl=below detection limit (0), adl=above detection limit (highest   
                         standard)", 
                         "Carbon dioxide (CO2) concentration", 
                         "Carbon dioxide (CO2) flag (bdl=below detection limit (0), adl=above detection limit (highest   
                         standard)", 
                         "Notes generated during field sample collection", 
                         "Notes generated during sample analysis in the lab") 

metadat$Units <-      c( "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA",
                         "NA",
                         "NA",
                         "m", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "ppm", 
                         "NA", 
                         "ppm", 
                         "NA", 
                         "NA", 
                         "NA")

metadat$Instrument_ifapplicable <-  c(
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA",
                         "NA",
                         "NA",
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "NA", 
                         "Varian 450 GC (Agilent Technologies)", 
                         "NA", 
                         "Varian 450 GC (Agilent Technologies)", 
                         "NA", 
                         "NA", 
                         "NA")


```

## Write out collated file 
```{r, include=FALSE}

##Collated data
write.csv(final_dat, "COMPASS_TEMPEST_TGW_AllData.csv")

##Metadata
write.csv(metadat, "COMPASS_TEMPEST_TGW_Metadata.csv")

```

## Plot each Sapflux Tree's Seasonal Trends
```{r, echo=FALSE}

final_dat$Sapflux_ID <- as.factor(final_dat$Sapflux_ID)
final_dat$Month_Num <- match(final_dat$Month, month.name)
final_dat <- final_dat %>% 
  mutate(Month = fct_reorder(Month, Month_Num)) %>% 
  mutate(CH4_ppm = as.numeric(CH4_ppm))  %>% 
  mutate(CO2_ppm = as.numeric(CO2_ppm))

#just plot all the data 
all_trees_CH4 <- ggplot(data=final_dat, aes(x=Month, y=as.numeric(CH4_ppm), color=Sapflux_ID))+
  geom_point()+ 
  facet_grid(~Plot, scales="free")+ 
  labs(y="CH4 (ppm)") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

all_trees_CH4

#just plot all the data 
all_trees_CO2 <- ggplot(data=final_dat, aes(x=Month, y=as.numeric(CO2_ppm), color=Sapflux_ID))+
  geom_point()+ 
  facet_grid(~Plot, scales="free")+ 
  labs(y="CO2 (ppm)") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

all_trees_CO2

```

##Plot boxplot summaries of CH4 and CO2 by Plot
```{r, echo=FALSE}

TGAS_avg <- final_dat %>%
  group_by(Plot, Sample_Grid) %>%
   dplyr::summarize(mean_CH4 = mean(CH4_ppm, na.rm=TRUE), 
            sd_CH4 = sd(CH4_ppm, na.rm=TRUE), 
            mean_CO2 = mean(CO2_ppm, na.rm=TRUE), 
            sd_CO2 = sd(CO2_ppm, na.rm=TRUE))


boxplot_CH4 <- ggplot(data=TGAS_avg, aes(x=Plot, y=mean_CH4, fill=Plot))+
  geom_boxplot()+ 
  labs(y="CH4 (ppm)")+
  theme_bw()

print(boxplot_CH4)


boxplot_CO2 <- ggplot(data=TGAS_avg, aes(x=Plot, y=mean_CO2, fill=Plot))+
  geom_boxplot()+ 
  labs(y="CO2 (ppm)")+
  theme_bw()

print(boxplot_CO2)


```

##Crate summary table of CH4/CO2 means by Plot
```{r, echo=FALSE}

TGAS_avg1 <- final_dat %>%
  group_by(Plot) %>%
   dplyr::summarize(mean_CH4 = mean(CH4_ppm, na.rm=TRUE), 
            sd_CH4 = sd(CH4_ppm, na.rm=TRUE), 
            mean_CO2 = mean(CO2_ppm, na.rm=TRUE), 
            sd_CO2 = sd(CO2_ppm, na.rm=TRUE))



kable(
  TGAS_avg1[, c("Plot", 	"mean_CH4", "sd_CH4",	"mean_CO2", "sd_CO2")],  
  col.names = c("Plot", "Mean [CH4] ppm", "std dev. CH4", "Mean [CO2] ppm", "std dev. CO2"),
  format = "latex",
  booktabs = TRUE,
  caption = "Summary Statistics by Site, and Zone") %>%
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%   # header
  column_spec(1, bold = TRUE)



# #Make a PNG from the tables for the READ ME 
# df_table <- TGAS_avg1[, c("Plot", 	"mean_CH4", "sd_CH4",	"mean_CO2", "sd_CO2")]
# 
# colnames(df_table) <- c("Plot", "Mean [CH4] ppm", "std dev. CH4", "Mean [CO2] ppm", "std dev. CO2")
# 
# tbl <- df_table %>%
#   gt() %>%
#   tab_header(
#     title = "Summary by Plot"
#   ) %>%
#   fmt_number(
#     columns = c(`Mean [CH4] ppm`, `std dev. CH4`, `Mean [CO2] ppm`, `std dev. CO2`),
#     decimals = 2
#   ) %>%
#   tab_style(
#     style = cell_text(weight = "bold"),
#     locations = cells_column_labels(everything())
#   )
# 
# tbl <- tbl %>%
#   cols_align(
#     align = "center",
#     columns = everything()
#   )
# 
# gtsave(tbl, "Tree_GHG_SumStats.png")



```


## Visualize Data by Plot   
```{r Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=20, fig.height=10, dpi = 300}

data_plotting <- final_dat %>% 
  mutate(Plot = factor(Plot, levels = c("Control", "Freshwater", "Saltwater")))

#Plot data and change colors based on Zone:
viz_ch4_plot <- ggplot(data_plotting, aes(x = as.character(Year), y = CH4_ppm, fill = Plot)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ Plot, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "CH4 (ppm)", title = "Samples: CH4") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 28))

# viz_ch4_plot


viz_co2_plot <-  ggplot(data_plotting, aes(x = as.character(Year), y = CO2_ppm, fill = Plot)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_wrap( ~ Plot, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "CO2 (ppm)", title = "Samples: CO2") +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1))+ 
  theme(text = element_text(size = 28))

# viz_co2_plot

ggarrange(viz_ch4_plot, viz_co2_plot, nrow=2, ncol=1)


```



## Summarized data for Site and Zone 
```{r Summarize and Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=22, fig.height=15, dpi = 300}

##Make Date Column
data_plotting$Date <- as.Date(paste(data_plotting$Year, data_plotting$Month, data_plotting$Day, sep = "-"), format = "%Y-%B-%d")

##Summarize TGAS by Date, Site, and Zone
TGAS_avg <- data_plotting %>%
  group_by(Date, Plot) %>%
   dplyr::summarize(mean_ch4 = mean(CH4_ppm, na.rm=TRUE), 
            sd_ch4 = sd(CH4_ppm, na.rm=TRUE), 
            mean_co2 = mean(CO2_ppm, na.rm=TRUE), 
            sd_co2 = sd(CO2_ppm, na.rm=TRUE))

# Define the specific date breaks you want to display
# specific_breaks <- as.Date(c("2022-01-01", "2023-01-01", "2024-01-01"))

viz_ch4_plot_avgs <- ggplot(TGAS_avg, aes(x = Date, y = mean_ch4, color = Plot)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_ch4 - sd_ch4, ymax = mean_ch4 + sd_ch4), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ Plot, scales="free_y", nrow = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "CH4 (ppm)", title = "Samples: CH4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 28))+
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y", expand = expansion(mult = c(0.18, 0.05)))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(size = 20))

# viz_ch4_plot_avgs


viz_co2_plot_avgs <- ggplot(TGAS_avg, aes(x = Date, y = mean_co2, color = Plot)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_co2 - sd_co2, ymax = mean_co2 + sd_co2), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ Plot, scales="free_y", nrow = 1) +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = " ", y = "CO2 (ppm)", title = "Samples: CO2") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 28))+
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y", expand = expansion(mult = c(0.18, 0.05)))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
  theme(legend.position = "none")+ 
  theme(axis.text.x = element_text(size = 20))

# viz_co2_plot_avgs

ggarrange(viz_ch4_plot_avgs, viz_co2_plot_avgs, nrow=2, ncol=1)

```


## Boxplot by species
```{r Boxplot by depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}


data_plotting_spp <- subset(data_plotting, !is.na(Tree_Species))

#Plot data and change colors based on Zone:
viz_co2_spp <-  ggplot(data_plotting_spp, aes(x = Tree_Species, y = CO2_ppm, fill = Plot)) +
  geom_boxplot() +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = "Tree Species", y = "CO2 (ppm)", title = "") + 
  theme(text = element_text(size = 30))+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

viz_co2_spp

#Plot data and change colors based on Zone:
viz_ch4_spp <-  ggplot(data_plotting_spp, aes(x = Tree_Species, y = CH4_ppm, fill = Plot)) +
  geom_boxplot() +
  scale_color_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  )) +
  theme_classic() +
  labs(x = "ree Species", y = "CH4 (ppm)", title = "") + 
  theme(text = element_text(size = 30))+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 

viz_ch4_spp

```


## Boxplot by Date
```{r Boxplot by Date, echo=FALSE, warning=FALSE, message = FALSE, fig.width=25, fig.height=10, dpi = 300}

data_plotting <- data_plotting %>% 
  mutate(Date_boxplot = gsub("-", "", Date))

##Summarize CO2 and CH4 by Date, Site, and Zone
boxplot_date_ch4 <- ggplot(data_plotting, aes(x = Date_boxplot, y = CH4_ppm, fill = Plot)) +
  geom_boxplot() + 
  facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CH4 (ppm)", title = "TEMPEST Tree Gas Well: CH4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

boxplot_date_ch4


boxplot_date_co2 <- ggplot(data_plotting, aes(x = Date_boxplot, y = CO2_ppm, fill = Plot)) +
  geom_boxplot() +
  facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CO2 (ppm)", title = "TEMPEST Tree Gas Well: CO2") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

boxplot_date_co2

```

## Boxplot by Grid
```{r Boxplot by Grid, echo=FALSE, warning=FALSE, message = FALSE, fig.width=25, fig.height=10, dpi = 300}

data_plotting_grid <- subset(data_plotting, !is.na(Sample_Grid))

##Summarize CO2 and CH4 by Date, Site, and Zone
boxplot_grid_ch4 <- ggplot(data_plotting_grid, aes(x = Sample_Grid, y = CH4_ppm, fill = Plot)) +
  geom_boxplot() + 
  facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CH4 (ppm)", title = "TEMPEST Tree Gas Well: CH4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

boxplot_grid_ch4


boxplot_grid_co2 <- ggplot(data_plotting_grid, aes(x = Sample_Grid, y = CO2_ppm, fill = Plot)) +
  geom_boxplot() +
  facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CO2 (ppm)", title = "TEMPEST Tree Gas Well: CO2") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

boxplot_grid_co2

```


## Boxplot by Tree
```{r Boxplot by Tree, echo=FALSE, warning=FALSE, message = FALSE, fig.width=25, fig.height=10, dpi = 300}

data_plotting_tree <- subset(data_plotting, !is.na(Sapflux_ID))

##Summarize CO2 and CH4 by Date, Site, and Zone
boxplot_tree_ch4 <- ggplot(data_plotting_tree, aes(x = Sapflux_ID, y = CH4_ppm, fill = Sample_Grid)) +
  geom_boxplot() + 
  facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  # scale_fill_manual(values = c(
  #   "Control" = "springgreen2",
  #   "Freshwater" = "cyan2",
  #   "Saltwater" = "violetred2"
  # ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CH4 (ppm)", title = "TEMPEST Tree Gas Well: CH4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm")) 
  # theme(legend.position = "none")

boxplot_tree_ch4


boxplot_tree_co2 <- ggplot(data_plotting_tree, aes(x = Sapflux_ID, y = CO2_ppm, fill = Sample_Grid)) +
  geom_boxplot() +
  facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  # scale_fill_manual(values = c(
  #   "Control" = "springgreen2",
  #   "Freshwater" = "cyan2",
  #   "Saltwater" = "violetred2"
  # ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CO2 (ppm)", title = "TEMPEST Tree Gas Well: CO2") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm")) 
  # theme(legend.position = "none")

boxplot_tree_co2

```

## Boxplot by Date
```{r Overall Boxplot, echo=FALSE, warning=FALSE, message = FALSE, fig.width=25, fig.height=10, dpi = 300}

##Summarize CO2 and CH4 by Date, Site, and Zone
viz_ch4_boxplot <- ggplot(data_plotting, aes(x = Plot, y = CH4_ppm, fill = Plot)) +
  geom_boxplot() + 
  # facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CH4 (ppm)", title = "TEMPEST Tree Gas Well: CH4") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_ch4_boxplot


viz_co2_boxplot <- ggplot(data_plotting, aes(x = Plot, y = CO2_ppm, fill = Plot)) +
  geom_boxplot() +
  # facet_grid( ~ Plot, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  scale_fill_manual(values = c(
    "Control" = "springgreen2",
    "Freshwater" = "cyan2",
    "Saltwater" = "violetred2"
  ), labels = c("Control", "Freshwater", "Saltwater")) + 
  theme_classic() +
  labs(x = " ", y = "CO2 (ppm)", title = "TEMPEST Tree Gas Well: CO2") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_co2_boxplot

# ggarrange(viz_CH4_boxplot, viz_CO2_boxplot, nrow=2, ncol=1)

# Save as a PNG
png("COMPASS_TEMPEST_CH4_alldata.png", width = 2000, height = 1200, res = 100) 
print(viz_ch4_boxplot)  # For ggplot objects, use print()
dev.off() 

# Save as a PNG
png("COMPASS_TEMPEST_CO2_alldata.png", width = 2000, height = 1200, res = 100) 
print(viz_co2_boxplot)  # For ggplot objects, use print()
dev.off() 

```
