---
title: "Soil respiration variability and correlation across a wide range of temporal scales"
author: "Ben Bond-Lamberty"
date: "28 October 2019"
output: html_document
---

## Introduction

* Noodling

```{r setup, include=FALSE}
library(drake)
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(kableExtra)
library(lattice)
library(cowplot)  # 0.9.4
theme_set(theme_bw())
#theme_set(theme_cowplot())

N_HISTOGRAM_BINS <- 20
START_DATE <- "2017-10-01"
END_DATE <- "2099-08-01"

readd("con_licor_data") %>% 
  filter(Timestamp >= START_DATE, Timestamp < END_DATE) %>% 
  arrange(Timestamp, Port) ->
  cld

```

Data limits are `r START_DATE` to `r END_DATE`.


```{r filtering}
cld %>% 
  filter(R2 > 0.75, !is.na(Flux)) %>%  # probably a chamber closing problem otherwise
  group_by(week(Timestamp)) %>% 
  mutate(mad = abs(Flux - median(Flux,  na.rm = TRUE)) / mad(Flux, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(mad <= 4) %>% 
  ungroup() ->
  cld_clean
removed_cld <- nrow(cld) - nrow(cld_clean)
```

Continuous data: `r nrow (cld)` observations; `r removed_cld` (`r round(removed_cld / nrow(cld) * 100, 0)`%) removed because of chamber closure or other problems. Final data is `r nrow(cld_clean)` observations.

Have data for `r length(unique(yday(cld_clean$Timestamp)))` unique days of the year.

```{r missing-dates}
# Insert missing dates
cld_clean$Timestamp <- as.POSIXct(round(cld_clean$Timestamp, "hour"))
mints <- min(cld_clean$Timestamp)
maxts <- max(cld_clean$Timestamp)

cld_clean %>% 
  group_by(Timestamp, Port) %>%
  summarise(Flux = mean(Flux, na.rm = TRUE))->
  cld_clean_avg

expand.grid(Port = 1:8,
            Timestamp = seq(mints, maxts, by = "hour")) %>% 
  as_tibble() %>%  
  left_join(cld_clean_avg, by = c("Port", "Timestamp")) %>% 
  mutate(Year = year(Timestamp), Month = month(Timestamp),
         Yday = yday(Timestamp), Hour = hour(Timestamp)) %>% 
  select(Timestamp, Year, Month, Yday, Hour, Port, Flux) ->
  cld_clean_hourly
```

After missing-date insertion, have data for `r length(unique(cld_clean_hourly$Yday))` unique days of the year.

## Distribution and variability

```{r continuous-plot, fig.height=8}
cld_clean_hourly %>% 
  mutate(Year = as.factor(Year)) %>% 
  ggplot(aes(yday(Timestamp), Flux, color = Year)) + 
  geom_line(alpha = 0.75, na.rm = TRUE) + facet_grid(Port ~ .) +
  ggtitle("A. Continuous data")
```

```{r distributions}
ggplot(cld_clean, aes(x = Flux, color = factor(Port))) + geom_density() +
  ggtitle("B. Distribution")
```


```{r continuous-collar-cv}
cld_clean_hourly %>% 
  mutate(Year = as.factor(year(Timestamp)),
         Hour = hour(Timestamp)) %>% 
  group_by(Year, Yday, Hour) %>% 
  summarise(Flux_cv = sd(Flux) / mean(Flux)) %>% 
  ungroup() ->
  cld_cv

cld_cv %>% 
  ggplot(aes(Yday, Flux_cv, color = Year)) + 
  geom_point(alpha = 0.5, na.rm = TRUE) +
  ggtitle("C. Collar-to-collar CV")

cld_clean_hourly %>% 
  mutate(Month = month(Timestamp),
         Hour = hour(Timestamp)) %>% 
  group_by(Month, Hour) %>% 
  summarise(Flux_cv = sd(Flux, na.rm = TRUE) / mean(Flux, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(Hour, Flux_cv, color = Month, group = Month)) + 
  geom_line() +
  ggtitle("D. Collar-to-collar CV by month")
```

## Gapfilling

```{r gapfill}
# Compute s.d. of the log values, by port and month
cld_clean_hourly %>% 
  filter(Flux > 0) %>% 
  group_by(Month, Port) %>% 
  summarise(Flux_logsd = sd(log(Flux), na.rm = TRUE),
            Flux_sd = sd(Flux, na.rm = TRUE)) ->
  cld_sd

# What's the difference if we use rnorm versus rlnorm (following XXX)

cld_clean_hourly %>% 
  left_join(cld_sd, by = c("Month", "Port")) %>% 
  arrange(Timestamp) %>% 
  group_by(Port) %>% 
  mutate(Flux_approx = approx(Timestamp, Flux, xout = Timestamp, rule = 2)$y) %>% 
  ungroup() ->
  cld_clean_gapfill

cld_clean_gapfill %>% 
  mutate(Type = "rnorm",
         Flux_approx_noise = if_else(is.na(Flux),
                                     abs(rnorm(n(), Flux_approx, Flux_sd)),
                                     Flux)) ->
  cld_clean_gapfill_noise

cld_clean_gapfill %>% 
  mutate(Type = "rlnorm",
         Flux_approx_noise = if_else(is.na(Flux),
                                     rlnorm(n(), log(Flux_approx), Flux_logsd),
                                     Flux)) ->
  cld_clean_gapfill_lognoise


cld_clean_gapfill_all <- bind_rows(cld_clean_gapfill_noise, cld_clean_gapfill_lognoise)

cld_clean_gapfill_all %>% 
  filter(is.na(Flux)) %>% 
  spread(Type, Flux_approx_noise) %>% 
  ggplot(aes(rnorm, rlnorm)) + geom_point(alpha = 0.5) + geom_abline(color = "blue") + 
  facet_wrap(~Port, scales = "free") ->
  p
print(p)

# Compute annual mean flux for each port and gapfill type
cld_clean_gapfill_all %>% 
  group_by(Type, Port, Yday) %>% 
  summarise(Flux_approx_noise = mean(Flux_approx_noise, na.rm = TRUE)) %>% 
  summarise(Flux_approx_noise = mean(Flux_approx_noise, na.rm = TRUE)) %>% 
  group_by(Type) %>% 
  summarise_at("Flux_approx_noise", tibble::lst(mean, sd))

cld_clean_gapfill_all %>% 
  filter(Port == 2) %>% 
  ggplot(aes(Timestamp, Flux_approx_noise, color = Type)) +
  geom_point() +
  facet_grid(Port ~ .) ->
  p
print(p)
```



As we changing sampling frequency, how does our estimate of Rs_annual change?

```{r monte-carlo, cache=TRUE}
compute_frac_mean <- function(frac, df) {
  df %>% 
    sample_frac(frac) %>% 
    pull(Flux_approx_noise) %>% 
    mean()
}

# This isn't what we want
# Want to compute the time length of the series; sample some fraction;
# and then adjust to an annual (or monthly or whatever) rate
# We want the "sample per year" variable to go to 365 * 24

set.seed(12345)
N_MC <- 100
cld_clean_gapfill_noise %>% 
  group_by(Port) %>% 
  summarise(meanflux = mean(Flux_approx, na.rm = TRUE)) ->
  meanflux

results <- list()
for(p in unique(cld_clean_gapfill_noise$Port)) {
  x <- filter(cld_clean_gapfill_noise, Port == p, !is.na(Flux_approx_noise))
  results[[p]] <- tibble(Port = p,
                         frac = rep(c(0.001, seq(0.01, 1, by = 0.05), 1.0), times = N_MC),
                         N_per_month = 365.25 * 24 / 12 * frac,
                         flux = sapply(frac, compute_frac_mean, x))
}
bind_rows(results) %>% 
  left_join(meanflux, by = "Port") %>% 
  mutate(flux_error = (flux - meanflux) / meanflux) %>% 
  group_by(Port, N_per_month, frac) %>% 
  summarise(n = n(), 
            flux_mean = mean(flux), 
            flux_sd = sd(flux),
            flux_error_mean = mean(flux_error),
            flux_error_sd = sd(flux_error)) ->
  y

ggplot(y, aes(N_per_month, flux_error_mean)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = flux_error_mean - flux_error_sd, 
                  ymax = flux_error_mean + flux_error_sd), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = c(1, 4, 10, 30, 720)) +
  coord_trans(x = "log10") + 
  facet_wrap(~Port)
```



## Session

```{r, echo=FALSE}
sessionInfo()
```
