---
title: "Seconds to decades: soil respiration variability and sensitivity across temporal scales"
author: "Ben Bond-Lamberty"
date: "11/17/2018"
output: html_document
---

```{r setup, include=FALSE}
# Much of this was presented at AGU 2018
library(drake)
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
theme_set(theme_bw())
library(lubridate)
library(kableExtra)
library(lattice)

library(mblm)  # for Theil-Sen robust trend test

licor_data <- readd("licor_data")

plot_dir <- "~/Desktop/agu_slides/"
save_agu_plot <- function(fn, plot = last_plot(), ...) {
  if(!dir.exists(plot_dir)) dir.create(plot_dir)
  suppressMessages(ggsave(file.path(plot_dir, fn), plot, ...))
}

fuzz <- function(x, error) {
  x * rnorm(length(x), mean = 1, sd = error)
}
```

## Seconds to minutes

How many times do we need to sample? Uses multiple Licor observations versus 1, 2, 3...

```{r howmanytimes, echo = FALSE}
licor_data %>% 
  filter(Flux < 10) %>% 
  group_by(Date, Group, Collar) %>%
  summarize(n = n(), mean_gt_2 = mean(Flux), mean_1 = Flux[1], mean_2 = mean(Flux[1:2])) %>% 
  filter(n >= 3) -> 
  lds

lds %>% 
  gather(variable, value, mean_1, mean_2) ->
  lds_plot

ggplot(lds_plot, aes(x = mean_gt_2, y = value, color = variable)) + 
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(method = "lm", linetype = 2) +
  geom_point() + 
  labs(x = "Mean flux of all (>=3) measurements", 
       y = "Mean flux of first 1 or 2 measurements")

lds %>% 
  gather(variable, value, mean_gt_2, mean_1, mean_2) %>% 
  mutate(variable = factor(variable, levels = c("mean_gt_2", "mean_2", "mean_1"))) ->
  lds_stats
m <- lm(value ~ variable, data = lds_stats)
summary(m)
TukeyHSD(aov(m))
```


* **Minutes to hours to days**: how frequently do we need to sample? Uses GCREW continuous data. How frequently do we need to sample?
* **Days to months**: how frequently do we need to sample for annual flux? Uses SERC survey data.
* **Years**: how much coverage is needed for robust annual estimate? This uses `Annual_coverage` field of SRDB. Not clear about this one.
* **Error introduced by RH versus RS**. Uses SRDB. Not sure about this. Hard.
* - our Nature anlysis assumed no measurement error. How could this be estimated?
* Random measurement error: "Random errors show up as different results for ostensibly the same repeated measurement. They can be estimated by comparing multiple measurements, and reduced by averaging multiple measurements." https://en.wikipedia.org/wiki/Observational_error#Random_errors_versus_systematic_errors
* - _continuous_ data: variability between measurement 1 and measurement 2
* - _survey_ (discontinuous) data: variability between m1 and m2 PLUS sampling error (sampling from continuous annual data)
* **Years to decades**: how well can we detect trends? This is Nature paper SI work.


### What's the SRDB interannual variability?

```{r srdb, echo=FALSE}
srdb <- readd("srdb")
srdb$Rs_interann_cv <- with(srdb, Rs_interann_err / Rs_annual)
median_interann_cv <- median(srdb$Rs_interann_cv, na.rm = TRUE)

srdb %>% 
  filter(!is.na(Rs_interann_cv)) %>% 
  ggplot(aes(x = Rs_interann_cv)) + geom_histogram(bins = 20) +
  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
  geom_vline(xintercept = median_interann_cv, color = "red") +
  ylab("Count") + xlab("CV between successive years")
save_agu_plot("srdb_cv.png")

# Site records
srdb %>% 
  filter(!is.na(Longitude), !is.na(Rs_annual), Manipulation == "None") %>% 
  mutate(lon = round(Longitude, 2), lat = round(Latitude, 2)) ->
  srdb

srdb %>% 
  # Compute the longest records
  group_by(lon, lat, Leaf_habit, Ecosystem_type) %>% 
  summarise(Years = length(unique(Study_midyear))) %>% ungroup %>% 
  arrange(desc(Years)) %>% 
  top_n(10) %>% 
  left_join(select(srdb, Site_name, lon, lat, Ecosystem_type, Study_midyear, YearsOfData),
            by = c("lon", "lat", "Ecosystem_type")) %>% 
  mutate(Site_name = substr(Site_name, 1, 10)) ->
  site_smry

ggplot(site_smry, aes(Study_midyear, Site_name, 
                      color = paste(Leaf_habit, Ecosystem_type))) + 
  geom_point() + xlab("Year") + ylab("") + scale_color_discrete("")
save_agu_plot("site_records.png", width = 8, height = 4)
```


### What's the CV between survey measurements 1 and 2?

```{r cv12, echo = FALSE}
licor_data %>% 
  group_by(Date, Group, Collar) %>%
  summarise(n = n(), meanFlux = mean(Flux), cv = sd(Flux) / mean(Flux)) %>% 
  filter(n == 2) ->
  meas_error_1

median_error <- median(meas_error_1$cv)
ggplot(meas_error_1, aes(x = cv)) + geom_histogram(bins = 20) +
  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
  geom_vline(xintercept = median_error, color = "red") +
  ylab("Count") + xlab("CV between successive IRGA measurements")
save_agu_plot("licor12_cv.png")
```

OK, so the median measurement error here is ~`r round(median_error * 100, 0)`% for `r nrow(meas_error_1)` observations of fluxes between `r round(min(meas_error_1$meanFlux), 2)` and `r round(max(meas_error_1$meanFlux), 2)` µmol/m2/s.

### Nature paper analysis

```{r hashimoto, echo=FALSE}
library(ncdf4)
# Downloaded August 25, 2017 from http://cse.ffpri.affrc.go.jp/shojih/data/index.html
ncfiles <- c("~/Data/Hashimoto/RH_yr_Hashimoto2015.nc",
             "~/Data/Hashimoto/RS_yr_Hashimoto2015.nc")

nc <- nc_open(ncfiles[1])
# These annual data start in 1901; extract 1990-2012
co2 <- ncvar_get(nc, "co2", start = c(1, 1, 1, 90), count = c(-1, -1, 1, 23))
nc_close(nc)


png(file.path(plot_dir, "co2.png"))
lattice::levelplot(co2[,,1])
dev.off()

co2 <- co2[400:600, 220:360,]  # punch a hole for testing: North America
#co2 <- co2[500:540, 320:360,]  # punch a hole for testing: part of North America

do_fitting <- function(co2) {
  
  f <- function(rh) { 
    df <- data.frame(x = seq_along(rh), y = rh)
    tryCatch(lm(y ~ x, data = df), error = function(e) NA)
  }
  
  # Fit linear model to each grid cell (this is slow)
  mods <- apply(co2, c(1, 2), FUN = f)  # slow
  
  # Extract slopes
  slopes <- apply(mods, c(1, 2), FUN = function(x) 
    if(!is.na(x)) x[[1]]$coefficients[["x"]] else NA)
  slopes <- matrix(slopes, nrow = nrow(mods), ncol = ncol(mods))
  
  # Extract slope p-values
  signif <- apply(mods, c(1, 2), FUN = function(x) 
    if(!is.na(x)) summary(x[[1]])$coefficients["x", "Pr(>|t|)"] else NA)
  signif <- matrix(signif, nrow = nrow(mods), ncol = ncol(mods))
  
  return(list(slopes = slopes, signif = signif))
}

out <- do_fitting(co2)
summary(as.vector(out$slopes))

png(file.path(plot_dir, "co2-slopes.png"))
lattice::levelplot(out$slopes > 0)
dev.off()
png(file.path(plot_dir, "co2-signif.png"))
lattice::levelplot(out$signif < 0.05)
dev.off()
hist(out$signif)

ncells <- sum(!is.na(out$slopes))
pos_slope <- sum(out$slopes > 0, na.rm = TRUE)
signif_pos_slope <- sum(out$slopes > 0 & out$signif < 0.05, na.rm = TRUE)

lat_weight <- abs(cos(seq(-pi/2, pi/2, length.out = nrow(out$slopes))))
ncells_areawt <- sum(lat_weight * ncol(out$slopes))
pos_slope_areawt <- sum(out$slopes > 0 * lat_weight, na.rm = TRUE)
signif_pos_slope_areawt <- sum(out$slopes > 0 & out$signif < 0.05 * lat_weight, na.rm = TRUE)
```

Total cells = `r ncells`.

Cells with positive slope = `r pos_slope` or `r round(pos_slope / ncells * 100, 0)`%.

Cells with _significant_ positive slope = `r signif_pos_slope` or `r round(signif_pos_slope / ncells * 100, 0)`%.

Area with positive slope = `r round(pos_slope_areawt / ncells_areawt * 100, 0)`%.

Area with _significant_ positive slope = `r round(signif_pos_slope_areawt / ncells_areawt * 100, 0)`%.

### Re-do analysis with assumed error rate

```{r fuzz, echo=FALSE}

co2_fuzz <- fuzz(co2, error = median_error)
out <- do_fitting(co2_fuzz)

png(file.path(plot_dir, "co2-fuzz-slopes.png"))
lattice::levelplot(out$slopes > 0)
dev.off()
png(file.path(plot_dir, "co2-fuzz-signif.png"))
lattice::levelplot(out$signif < 0.05)
dev.off()
hist(out$signif)

ncells <- sum(!is.na(out$slopes))
pos_slope <- sum(out$slopes > 0, na.rm = TRUE)
signif_pos_slope <- sum(out$slopes > 0 & out$signif < 0.05, na.rm = TRUE)

lat_weight <- abs(cos(seq(-pi/2, pi/2, length.out = nrow(out$slopes))))
ncells_areawt <- sum(lat_weight * ncol(out$slopes))
pos_slope_areawt <- sum(out$slopes > 0 * lat_weight, na.rm = TRUE)
signif_pos_slope_areawt <- sum(out$slopes > 0 & out$signif < 0.05 * lat_weight, na.rm = TRUE)

# Convert to a data frame for ggplot2 plotting
ro = nrow(co2_fuzz)
co = ncol(co2_fuzz)
yr = dim(co2_fuzz)[3]
co2_fuzz_df <- tibble(
  flux = as.vector(co2_fuzz),
  lat = rep(seq_len(ro), times = co * yr),
  lon = rep(rep(seq_len(co), each = ro), times = yr),
  year = rep(seq_len(yr), each = ro * co),
  p = rep(as.vector(out$signif), times = yr)
)

co2_fuzz_df %>% 
  filter(!is.na(p)) %>% 
  # pick a subset of grid cells for a readable plot
  distinct(lon, lat) %>% 
  sample_n(250) %>% 
  left_join(co2_fuzz_df, by = c("lon", "lat")) ->
  co2_fuzz_subsampled

co2_fuzz_subsampled %>% 
  ggplot(aes(year + 1990, flux, group = paste(lat, lon))) + 
  geom_line(color = "lightgrey") +
  xlab("Year") + ylab("Flux (gC/m2/yr)") +
  geom_line(data = filter(co2_fuzz_subsampled, p < 0.05), color = "red", alpha = I(0.5))
save_agu_plot("fuzz_over_time.png")
```

Total cells = `r ncells`.

Cells with _significant_ positive slope (observations with `r round(median_error * 100, 0)`% measurement error) = `r signif_pos_slope` or `r round(signif_pos_slope / ncells * 100, 0)`%.

Area with _significant_ positive slope = `r round(signif_pos_slope_areawt / ncells_areawt * 100, 0)`%.


# Next steps

Next: make a nice graph of change over time 
using a subset of data for readability
Convert array to data frame and plot rs versus time
with a line for each grid cell

# Simple: when would expect to see significance?

* We'd like to do this once for perfect data
* Once for data + interannual variability
* Once for data + iav + observational error

```{r simple, echo=FALSE}
set.seed(1234)

trend_emergence <- function(rd, theilsen = F) {
  Year <- seq_len(length(rd))
  trend_p <- rep(NA, length(rd))
  for(i in seq_along(trend_p)) {
    if(i > 2) {
      if(theilsen) {
        df <- tibble(Year = Year[1:i], rd = rd[1:i])
        suppressWarnings(m <- mblm::mblm(rd ~ Year, data = df))  # mblm doesn't like form below
      } else {
        m <- suppressWarnings(lm(rd[1:i] ~ Year[1:i]))
      }
      # Extract 2nd row (Year) and 4th column (Pr>[t] or Pr>|V|)
      trend_p[i] <- summary(m)$coefficients[2, 4]
    }
  }
  trend_p
}


# Temperature has risen 0.9 C in 40 years, more or less
dTdt <- round(0.9 / 40.0, 3)
q10 <- 2
R0 = 1.0
respdata <- tibble(Year = 1:100,
                   Temp = dTdt * Year,
                   Resp = R0 * q10 ^ (Temp / 10),
                   # This is interannual variability
                   Resp_iav = fuzz(Resp, 0.098),  # this is SRDB Rs_interannual_err
                   Resp_fuzz = fuzz(Resp_iav, median_error))

# Make a nice plot--first with ideal curve, then IAV, then observations
p <- ggplot(respdata, aes(Year, Resp)) + 
  geom_point(color = "grey") + 
  ylab("Respiration") + coord_cartesian(ylim = c(0.75, 1.5)) + 
  annotate("text", 10, 1.4, label = paste("Q10 =", q10)) + 
  annotate("text", 10, 1.3, label = paste("dT/dt =", dTdt))
save_agu_plot("01.png")
p <- p + geom_point(aes(y = Resp_iav))
save_agu_plot("02.png")
p <- p + geom_errorbar(aes(ymin = Resp_iav - Resp_iav * median_error,
                           ymax = Resp_iav + Resp_iav * median_error))
save_agu_plot("03.png")

save_agu_plot("04.png",
              p + geom_line(aes(y = Resp), color = "red", size = 2))

save_agu_plot("05.png",
              p + geom_line(aes(y = Resp), color = "pink", size = 2) +
                geom_line(aes(y = Resp_iav), color = "red", size = 2))

save_agu_plot("06.png",
              p + geom_line(aes(y = Resp), color = "pink", size = 2) +
                geom_line(aes(y = Resp_iav), color = "pink", size = 2) +
                geom_line(aes(y = Resp_fuzz), color = "red", size = 2))


do_sim <- function(i, respdata, error = 0.0) {
  # This is observational error
  respdata$Resp_fuzz <- fuzz(respdata$Resp_iav, error)
  respdata$trend_p <- trend_emergence(respdata$Resp_fuzz)
  respdata
}

results <- list()
library(parallel)
n_sims <- 100
results <- mclapply(seq_len(n_sims), do_sim, respdata, error = median_error)

results %>% 
  bind_rows %>% 
  group_by(Year) %>% 
  summarise(n = n(), 
            Temp = mean(Temp), 
            Resp = mean(Resp),
            Resp_iav_sd = sd(Resp_iav),
            Resp_iav = mean(Resp_iav),
            Resp_fuzz_sd = sd(Resp_fuzz),          
            Resp_fuzz = mean(Resp_fuzz), 
            trend_p_sd = sd(trend_p), 
            trend_p = mean(trend_p)) %>% 
  filter(!is.na(trend_p)) ->
  results_summary

p <- ggplot(results_summary, aes(Year, trend_p, color = trend_p < 0.05)) +
  geom_point() +
  geom_line(aes(y = Resp_fuzz)) +
  geom_line(aes(y = Resp), color = "grey") +
  geom_ribbon(aes(ymin = Resp_fuzz - Resp_fuzz_sd, 
                  ymax = Resp_fuzz + Resp_fuzz_sd, 
                  fill = trend_p < 0.05), color = NA, alpha = I(0.35)) +
  guides(color = FALSE, fill = FALSE) +
  annotate("text", 10, 1.5, label = paste("N =", n_sims)) +
  ylab("Theil-sen p-value   ///   Respiration")
print(p)
save_agu_plot("simple_sim.png")
```
