---
title: "Sapflow QA / QC"
author: "Ben Bond-Lamberty and Stephanie Pennington"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Load packages
TESTING <- TRUE
VOLTAGE_RANGE <- c(0.2, 0.6)  # expected 'good' range for probe voltage diff

# Daily cycle settings
AMPLITUDE_RANGE <- c(-0.01, -0.2)

# This list holds the test results. The name of each list entry is of the form
# "WARN.<test_name" or "ERROR.<test_name>".
# Every entry is a tibble/data frame, with the following required columns:
# * Tree_Code
# * Port
# * failures - an integer
DAILY_TESTS <- list()

library(readr)
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
theme_set(theme_minimal())

# Global chunk options
knitr:: opts_chunk$set(echo = FALSE)

# Function setup
read_sapflow <- function(filename) {
  sdat <- readLines(filename)
  sdat <- sdat[-3:-4] # remove lines 3 and 4 with unneeded information
  
  # parse line one to extract logger name
  pnnl_x <- gregexpr("PNNL_", sdat[1])[[1]][1]
  logger_name <- substr(sdat[1], start = pnnl_x, stop = pnnl_x + 6)
  
  read_csv(sdat, skip = 1) %>% 
    mutate(Logger = logger_name)
}

# Generate list of 'current' sapflow files 
s_files <- list.files(path = "../Data/sapflow/", pattern = "PNNL", full.names = TRUE)
ports <- read_csv("../Design/ports.csv", col_types = "ldcdc")
```

```{r process-data}
# Read in data we're reporting on, reshape and reformat for analysis, etc.

if(TESTING) {
  message("TESTING!")
  sapflow <- read_csv("test_data/sapflow_test_20210601.csv")
} else {
  # Read in data
  lapply(s_files, read_sapflow) %>%  
    bind_rows() %>% 
    # extract number form former col name "DiffVolt_Avg(1)" would become "1"
    # join with ports dataframe and bring in tree codes
    pivot_longer(cols = `DiffVolt_Avg(1)`:`DiffVolt_Avg(6)`, 
                 names_to = "Port", values_to = "Value") %>% 
    rename(Timestamp = TIMESTAMP, 
           Record = RECORD) %>% 
    mutate(Timestamp = ymd_hms(Timestamp, tz = "EST"), 
           Port = parse_number(Port), 
           Logger = parse_number(Logger)) -> sf_raw
  
  sf_raw %>% 
    left_join(ports, by = c("Logger", "Port")) %>% 
    select(Timestamp, Record, BattV_Avg, Port, Value, Logger, Tree_Code, Species) -> sapflow
  
  nomatch <- anti_join(sf_raw, ports, by = c("Logger", "Port"))
  if(nrow(nomatch) > 0) {
    warning("There were logger/port combinations that I couldn't find in ports.csv:")
    nomatch %>% 
      distinct(Logger, Port) %>% 
      print()
  }
}

# Add some extra time information
sapflow %>% 
  mutate(Date = as.Date(Timestamp),
         Hour = hour(Timestamp) + minute(Timestamp) / 60) -> 
  sapflow

# BBL-SP todo
# * test dataset should include data that don't have a logger/port match
# * test code above should read RAW test data ideally
# * check sapflow manual - confirm we should see high values at night, low during day 
# * duplicated data? May want to distinct() sapflow
```


```{r create-test-dataset, eval=FALSE}
# Create test dataset - not normally run
trees <- c("F16", "S15", "F2", "S6", "S2", "F13", "C10", "C11", "L4")

sapflow %>% 
  filter(Tree_Code %in% trees,
         Timestamp > "2021-04-25 12:30:00") -> test_df
```

What do we want to identify? h/t to SP's lovely test data here

OUT OF RANGE TEST
* values out of range {0.2, 0.6} (from manual)
* also: days with values outside of {0.2, 0.6}

DAILY CYCLE TEST
* days exhibiting a 'normal' cycle

TREND TEST (?)
* if data exhibiting trend over X days

I think each section (code chunk = test) produces a data frame with day, tree, error_code
ERROR_xxxxx
WARNING_xxxxx
NOTE_xxxxxx

It sticks this into a list

At the end of the script, we assemble all of these into a single summary data frame

# Data statistics

```{r, warning=FALSE}
library(skimr)
skim(sapflow)
```

# Tests

## Out of range test

This test flags data outside of the expected good voltage range: `r paste(VOLTAGE_RANGE, collapse = ", ")`.

```{r out-of-range-test}
# SCP
```

## Missing data test

This test flags missing (`NA`) data.

```{r missing-test}
# SCP
```

## Daily cycle test

This test identifies days that don't have a normal-looking daily cycle (here defined simply
as having afternoon values that are lower than non-afternoon values; allowable range is
`r paste(AMPLITUDE_RANGE, collapse = ", ")`). 

```{r daily-cycle-test}
# Helper function
daily_cycle_amplitude <- function(hours, values) {
  stopifnot(length(hours) == length(values))
  stopifnot(!any(duplicated(hours)))
  
  # We need data coverage over the day to do this
  if(min(hours) > 4 && max(hours) < 20 || length(hours) < 50) {
    return(NA)
  }
  
  # Simple test: voltage during afternoon hours (based on known good data) should be lower
  # than others
  afternoon <- hours %in% 12:18
  mean(values[afternoon]) - mean(values[!afternoon])
}

# Compute 'amplitude' (kind of; the difference between afternoon and non-afternoon values)
sapflow %>% 
  group_by(Date, Tree_Code, Port) %>% 
  summarise(amplitude = daily_cycle_amplitude(Hour, Value),
            good_cycle = between(amplitude, min(AMPLITUDE_RANGE), max(AMPLITUDE_RANGE)),
            .groups = "drop") ->
  daily_amplitude_data

# Store results
daily_amplitude_data %>% 
  group_by(Tree_Code, Port) %>%
  summarise(failures = sum(!good_cycle, na.rm = TRUE), .groups = "drop") ->
  DAILY_TESTS[["WARN.daily_cycle"]]

sapflow %>% 
  left_join(daily_amplitude_data, by = c("Date", "Tree_Code", "Port")) ->
  sapflow

ggplot(sapflow, aes(Timestamp, Value, color = good_cycle)) + 
  geom_point() + 
  facet_wrap(~Tree_Code) + 
  coord_cartesian(ylim = c(0, 1))
```

## Trend test

```{r}
TREND_DAYS <- 7
```

This test looks at every `r TREND_DAYS` of data and tests whether there's a significant 
trend (using a Mann-Kendall test) in the daily mean.

```{r trend-test}
sapflow %>% 
  mutate(group = yday(Timestamp) %/% TREND_DAYS) ->
  sapflow

library(Kendall)

mk_no_trend <- function(x) {
  if(length(x) < 3) NA else  MannKendall(x)$sl > 0.05
}

sapflow %>% 
  group_by(Date, group, Tree_Code, Port) %>% 
  summarise(Value = mean(Value), .groups = "drop") %>% 
  group_by(group, Tree_Code, Port) %>% 
  summarise(no_trend = mk_no_trend(Value),
            .groups = "drop") ->
  weekly_trends

# Store results
weekly_trends %>% 
  group_by(Tree_Code, Port) %>%
  summarise(failures = sum(!no_trend, na.rm = TRUE), .groups = "drop") ->
  DAILY_TESTS[["WARN.trend"]]

sapflow %>% 
  left_join(weekly_trends, by = c("group", "Tree_Code", "Port")) ->
  sapflow

ggplot(sapflow, aes(Timestamp, Value, color = no_trend)) + 
  geom_point() + 
  facet_wrap(~Tree_Code) + 
  coord_cartesian(ylim = c(0, 1))
```

# Summary

## Daily test failure count:

```{r summary, warning=FALSE}
DAILY_TESTS %>% 
  bind_rows(.id = "test") %>% 
  separate(test, into = c("Test_type", "Test_name"), sep = "\\.") %>% 
  pivot_wider(names_from = "Test_name", values_from = "failures", values_fill = 0) ->
  daily_test_summary

library(DT)
datatable(daily_test_summary, 
          rownames = FALSE,
          filter = "top",
          options = list(pageLength = 10, scrollX = TRUE))
```

