---
title: "TEMPEST: Soil Gas Well"
author: "May 2025 Samples"
date: "`r Sys.Date()`"
output:  
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information
```{r run information, include=TRUE}
cat("Run Information: Brenden Blakley") #lets you know what section you're in
#set the run date & user name
sample_year <- "2025"
sample_month <- "May"
user <- "Brenden Blakley"
#identify the files you want to read in
#read in as a list to accommodate multiple runs in a month (.txt file(s))

GHG_files <- c("Raw Data/20260827_TMP2025_May.txt")

#Metadata file path
Metadata_path = "Metadata/TEMPEST_SGAS_202505_ALL_WORKING.csv"


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "GC_QAQC_Log.csv"
check_log_path <- "GC_Check_QAQC_Log.csv"


# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/TMP2025_May.csv"

##Standard Type ("Actual" or "Half")
 #LTREB, SMARTX, and GENX PW Samples use "Half", others should use "Actual"
Standard_Type <- "Actual"

#Did Some Samples require manual area retrieval? 
##List Samples and their areas for each analyte ("CH4", "CO2", or "N2O")
Manual_Areas <- data.frame(
  Sample = c("NA"),
  Area = c("NA"),
  Analyte = c("NA")
)

#Gas in Water Calculations
#Are All Samples Gas in Water? ("YES" or "NO")
GIW <- "NO"
#Are some samples Gas in Water? 
##List Sample IDs
GIW_Samples <- c()

# "TempC" column value in Celsius (equilibration temperature) and volume of gas (in liters) in syringe (not changing currently)
TempC = 25 #Equilibration Temperature (Celsius)

volume_g = 0.015 #volume of gas (in liters) in syringe when equilibrized
volume_w = 0.015 #volume of water (in liters) in syringe when equilibrized (should be equal to volume_g)

#record any notes about the run or anything other info here:
run_notes <- "Some samples misnamed - corrected"

#Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

cat(run_notes)
```


##Set Up

```{r setup, include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  fuzzyjoin,
  ggplot2,
  ggpubr,
  stringr,
  here,
  data.table,
  tidyverse,
  lubridate)

#add figure alignment for pdf output
knitr::opts_chunk$set(
  fig.align = 'left',
  echo = FALSE
)

##Low & High Standard Curve Cutoffs
low_std_crv_cutoff_ch4 <- 100
low_std_crv_cutoff_N2O <- 50.2
r2_cutoff <- 0.990

##Gas In Water Calculation
R <- 0.08206
CTK <- 273.15

##Detection Limits
#adl_cutoff <- 10000
#bdl_cutoff <- 1

##Check Standard Concentration
chks_flag = 60
blks_flag = 60
if(Standard_Type == "Actual"){
level_ch4 <- 500 #ppm
level_co2 <- 25000 #ppm
level_n2o <- 250} else {
level_ch4 <- 1000 #ppm
level_co2 <- 50000 #ppm
level_n2o <- 500 #ppm
}

##CH4 Retention Time Cutoffs
CH4_Low <- 1.98
CH4_High <- 2.08

##CO2 Retention Times Cutoffs
CO2_Low <- 2.62
CO2_High <- 2.72

##N2O Retention Times Cutoffs
N2O_Low <- 2.84
N2O_High <- 2.94

##Standard Concentrations (ppm)
CH4_LowTri <- 10
CO2_LowTri <- 2000
N2O_LowTri <- 5.05

CH4_MidTri <- 100
CO2_MidTri <- 20000
N2O_MidTri <- 50.2

CH4_HighTri <- 1000
CO2_HighTri <- 50000
N2O_HighTri <- 500

CH4_10K <- 10000
CO2_10K <- 0
N2O_10K <- 0

```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

Raw_Metadata <- read.csv(Metadata_path)

Clean_Metadata <- Raw_Metadata %>%
  dplyr::select(-Machine,
                -User,
                -Run_Date,
                -Dilution_Factor,
                -STD_Conc,
                -CO2_Area,
                -CH4_Area) %>% 
  filter(Sample_Type == "Unknown") %>%
  mutate(Sample_Name = paste(Sample_Year, Sample_Month, Sample_ID, sep = "_")) %>%
  select(-Sample_ID)
  


```

##Read In Data Files
```{r}

#Read in Nov Data
raw_list <- map(GHG_files, read.delim, header = FALSE, row.names = NULL, sep = ",", quote = "", col.names = paste0("V", 1:21))


# Combining Files
DataFile <- bind_rows(raw_list)


```

#Get Location of Each Sample in Main Data Frame
```{r}
Numbers <- c(which(DataFile$V1 == "[Header]"), (length(DataFile$V1)+1))
```

#Turn Each Sample into it's own Data Frame
```{r}
n_dfs <- (length(Numbers)-1)

df_list <- replicate(n_dfs, data.frame(), simplify = FALSE)

for(i in 1:n_dfs) {
  df_loop1 <- as.data.frame(DataFile[(Numbers[i]):(Numbers[i+1] - 1),])
  df_list[[i]] <- df_loop1
}
```


#Get Only Area Tables and Retrieve Info
```{r}

Sample_Data <- data.frame(Sample_Name = c(NA), Area1 = c(NA), RTime1 = c(NA), Area2 = c(NA), RTime2 = c(NA), Area3 = c(NA), RTime3 = c(NA), NumPeakFID = c(NA), NumPeakECD = c(NA), Run_Date = c(as.Date(NA)))

for(j in 1:(length(Numbers)-1)) {
  
#FID
  df_loop2 <- df_list[[j]] #get only 1 sameple
  peak_start <- which(df_loop2$V1 == "[Peak Table(Ch1)]") #find where peak data starts
  peak_end <- which(df_loop2$V1 == "[Peak Table(Ch2)]") #find where peak data ends
  sample <- df_loop2[which(df_loop2$V1 == "Sample Name"),]$V2 #get sample name
  df_peaks <- df_loop2[peak_start:(peak_end-1),] #subset data to get only peak info
  num_peaks <- df_peaks[which(df_peaks$V1 == "# of Peaks"),]$V2 #get number of found peaks (should be 2, possibly 1 if blank/low)
  table_start <- which(df_peaks$V1 == "Peak#") #find where table headers start
  area_col <- which(df_peaks[table_start,] == "Area") #get column of area
  rtime_col <- which(df_peaks[table_start,] == "R.Time") #get column of retention time
  if(num_peaks == 0) {peak1_area <- NA #get peak 1 info if present (should be CH4)
                      peak1_rtime <- NA} else { 
         peak1_area <- df_peaks[which(df_peaks$V1 == 1), area_col]
         peak1_rtime <- df_peaks[which(df_peaks$V1 == 1), rtime_col]}
  if(num_peaks <= 1) {peak2_area <- NA #get peak 2 info if present (should be CO2)
                      peak2_rtime <- NA} else { 
         peak2_area <- df_peaks[which(df_peaks$V1 == 2), area_col]
         peak2_rtime <- df_peaks[which(df_peaks$V1 == 2), rtime_col]}
  
#ECD
  ECD_peak_start <- which(df_loop2$V1 == "[Peak Table(Ch2)]") #Find ECD Peak Data start
  ECD_peak_end <- nrow(df_loop2) #get peak data end
  df_ecd_peaks <- df_loop2[ECD_peak_start:ECD_peak_end,] #get only ecd peak info
  ecd_num_peaks <- df_ecd_peaks[which(df_ecd_peaks$V1 == "# of Peaks"),]$V2  #get number of found peaks (should be 1, possibly 0 if blank)
  ecd_table_start <- which(df_ecd_peaks$V1 == "Peak#") #find ecd table headers
  ecd_area_col <- which(df_ecd_peaks[table_start,] == "Area") #get column of area
  ecd_rtime_col <- which(df_ecd_peaks[table_start,] == "R.Time") #get column of retention time
  if(ecd_num_peaks == 0) {peak3_area <- NA #get peak info if present
                         peak3_rtime <- NA } else {
                           peak3_area <-df_ecd_peaks[which(df_ecd_peaks$V1 == 1), ecd_area_col]
                           peak3_rtime <- df_ecd_peaks[which(df_ecd_peaks$V1 == 1), ecd_rtime_col]}
  
   sample_run_date <- as.Date(df_loop2$V2[which(df_loop2$V1 == "Acquired")[1]], format = "%m/%d/%Y %H:%M") #get the date sample was ran
##Assign Values
  Sample_Data[j,]$Sample_Name <- sample
  Sample_Data[j,]$Area1 <- peak1_area
  Sample_Data[j,]$RTime1 <- peak1_rtime
  Sample_Data[j,]$Area2 <- peak2_area
  Sample_Data[j,]$RTime2 <- peak2_rtime
  Sample_Data[j,]$Area3 <- peak3_area
  Sample_Data[j,]$RTime3 <- peak3_rtime
  Sample_Data[j,]$NumPeakFID <- num_peaks
  Sample_Data[j,]$NumPeakECD <- ecd_num_peaks
  Sample_Data[j,]$Run_Date <- sample_run_date
}

```


##QAQC Peak Count and Retention Time
```{r}
#Flag which samples have the wrong amount of peaks FID should have 2, ECD 1
PeakFlags <- Sample_Data %>% mutate(FID_Peak_Flag = ifelse(NumPeakFID == 0, "Incorrect - No Peaks", ifelse(NumPeakFID == 1, "Incorrect - One Peak", ifelse(NumPeakFID == 2, "Correct", "Incorrect - Too Many Peaks")))) %>% mutate(ECD_Peak_Flag = ifelse(NumPeakECD == 0, "Incorrect - No Peaks", ifelse(NumPeakECD == 1, "Correct", "Incorrect - Too Many Peaks")))

#Flag any retention times that are out of range for compound
RTimeFlags <- PeakFlags %>% mutate(CH4_RTime_Flag = ifelse(RTime1 >= CH4_Low & RTime1 <= CH4_High, "Correct - CH4 RTime", "Incorrect - Wrong RTime")) %>% mutate(CO2_RTime_Flag = ifelse(RTime2 >= CO2_Low & RTime2 <= CO2_High, "Correct - CO2 RTime", "Incorrect - Wrong RTime")) %>% mutate(N2O_RTime_Flag = ifelse(RTime3 >= N2O_Low & RTime3 <= N2O_High, "Correct - N2O RTime", "Incorrect - Wrong RTime" ))

#Display Flags
paste0("If there are too many peaks it's possible there is a problem with the sample or run and the chromatogram should be checked, if there are too few they may be low/blanks, verify that concentration is low or get area manually")
Flagged_FID_Peaks <- ifelse(RTimeFlags$FID_Peak_Flag %ilike% "Incorrect",paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$NumPeakFID), NA)
Flagged_FID_Peaks2 <- na.omit(Flagged_FID_Peaks)
paste("FID:",Flagged_FID_Peaks2)

Flagged_ECD_Peaks <- ifelse(RTimeFlags$ECD_Peak_Flag %ilike% "Incorrect",paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$NumPeakECD), NA)
Flagged_ECD_Peaks2 <- na.omit(Flagged_ECD_Peaks)
paste("ECD:", Flagged_ECD_Peaks2)

paste0("If RTime is wrong check that correct number of peaks were found, if not a peak may have been mislabeled, if peak number is correct then there may be a problem with the run, get area manually")
CH4_Flagged_RTimes <- ifelse(RTimeFlags$CH4_RTime_Flag %ilike% "Incorrect", paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$RTime1), NA)
CH4_Flagged_RTimes2 <- na.omit(CH4_Flagged_RTimes)
paste("CH4: ", CH4_Flagged_RTimes2)

CO2_Flagged_RTimes <- ifelse(RTimeFlags$CO2_RTime_Flag %ilike% "Incorrect", paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$RTime2), NA)
CO2_Flagged_RTimes2 <- na.omit(CO2_Flagged_RTimes)
paste("CO2: ", CO2_Flagged_RTimes2)

N2O_Flagged_RTimes <- ifelse(RTimeFlags$N2O_RTime_Flag %ilike% "Incorrect", paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$RTime3), NA)
N2O_Flagged_RTimes2 <- na.omit(N2O_Flagged_RTimes)
paste("N2O: ", N2O_Flagged_RTimes2)

Flagged_Data <- RTimeFlags

#Correct Missnamed Samples
Flagged_Data$Sample_Name[31] <- "ChkStd_HighTri_12mL_2"
Flagged_Data$Sample_Name[32] <- "Blank5"
Flagged_Data$Sample_Name[33] <- "2025_May_11"
Flagged_Data$Sample_Name[43] <- "ChkStd_HighTri_12mL_3"
Flagged_Data$Sample_Name[44] <- "Blank6"
Flagged_Data$Sample_Name[45] <- "2025_May_21"
Flagged_Data$Sample_Name[46] <- "2025_May_22"






```

##Format Data
```{r}
#Assign 0 to Blanks without peaks
Data_Zero <- Flagged_Data %>% mutate(Area1 =  ifelse(is.na(Area1) & Sample_Name %ilike% "Blank|0ppm", 0, Area1),
                                     Area2 = ifelse(is.na(Area2) & Sample_Name %ilike% "Blank|0ppm|10K", 0, Area2),
                                     Area3 = ifelse(is.na(Area3) & Sample_Name %ilike% "Blank|0ppm|10K", 0, Area3))

#Rename Columns
Data_RNM <- Data_Zero %>% rename(Sample_ID = Sample_Name,
                                    CH4_Area = Area1,
                                    CO2_Area = Area2,
                                    N2O_Area = Area3)

#Get Only Relevant Columns
Data_Sum <- Data_RNM %>% dplyr::select(Sample_ID,
                                       Run_Date,
                                       CH4_Area,
                                       CH4_RTime_Flag,
                                       CO2_Area,
                                       CO2_RTime_Flag,
                                       N2O_Area,
                                       N2O_RTime_Flag,
                                       FID_Peak_Flag,
                                       ECD_Peak_Flag)
```

##Add Standard Concentrations
```{r}

#Assign Source Standard Concentrations
Data_Stds <- Data_Sum %>% 
  mutate(CH4_STD = ifelse(Sample_ID %ilike% "HighTri", CH4_HighTri, 
                               ifelse(Sample_ID %ilike% "MidTri", CH4_MidTri, 
                                      ifelse(Sample_ID %ilike% "LowTri", CH4_LowTri, 
                                             ifelse(Sample_ID %ilike% "10K", CH4_10K, ifelse(Sample_ID %ilike% "0ppm", 0, NA)))))) %>% 
  mutate(CO2_STD = ifelse(Sample_ID %ilike% "HighTri", CO2_HighTri, 
                               ifelse(Sample_ID %ilike% "MidTri", CO2_MidTri, 
                                      ifelse(Sample_ID %ilike% "LowTri", CO2_LowTri, 
                                             ifelse(Sample_ID %ilike% "10K", CO2_10K, ifelse(Sample_ID %ilike% "0ppm", 0, NA)))))) %>% 
  mutate(N2O_STD = ifelse(Sample_ID %ilike% "HighTri", N2O_HighTri, 
                               ifelse(Sample_ID %ilike% "MidTri", N2O_MidTri, 
                                      ifelse(Sample_ID %ilike% "LowTri", N2O_LowTri, 
                                             ifelse(Sample_ID %ilike% "10K", N2O_10K, ifelse(Sample_ID %ilike% "0ppm", 0, NA)))))) %>%
  mutate(mL_STD = ifelse(Sample_ID %ilike% "Tri|10K", sub(".*_([0-9]+)mL.*", "\\1", Sample_ID), NA)) %>% #Get mL of source standard
  mutate(mL_STD = as.numeric(mL_STD), #Convert Values
         CH4_STD = as.numeric(CH4_STD),
         CO2_STD = as.numeric(CO2_STD),
         N2O_STD = as.numeric(N2O_STD)) %>% 
  mutate(Conc_CH4 = round(CH4_STD * mL_STD / 24, digits = 4), #Calculate Concentrations of Standards
         Conc_CO2 = round(CO2_STD * mL_STD / 24, digits = 4),
         Conc_N2O = round(N2O_STD * mL_STD / 24, digits = 4),) %>%
  mutate(Conc_CH4 = ifelse(Sample_ID %ilike% "0ppm", 0, Conc_CH4),
         Conc_CO2 = ifelse(Sample_ID %ilike% "0ppm", 0, Conc_CO2),
         Conc_N2O = ifelse(Sample_ID %ilike% "0ppm", 0, Conc_N2O),)

```

##Finalize Data
```{r}
raw <-Data_Stds %>% mutate(Standard = Standard_Type) %>%
  mutate(Conc_CH4 = ifelse(Standard == "Half", Conc_CH4 * 2, Conc_CH4), #Adjust standard concentrations for dilutions if necessary
  Conc_CO2 = ifelse(Standard == "Half", Conc_CO2 * 2, Conc_CO2),
   Conc_N2O = ifelse(Standard == "Half", Conc_N2O * 2, Conc_N2O)) %>% 
  mutate(Type = ifelse(Sample_ID %ilike% "Std|0ppm" & !(Sample_ID %ilike% "Chk|Check"), "Standard", 
                       ifelse(Sample_ID %ilike% "Lab", "Lab air", 
                              ifelse(Sample_ID %ilike% "Blank", "Blank", 
                                     ifelse(Sample_ID %ilike% "Chk|Check", "ChkStd", "Sample") )))) %>% #Add Type
  mutate(CH4_Area = as.numeric(CH4_Area),
         Conc_CH4 = as.numeric(Conc_CH4),
         CO2_Area = as.numeric(CO2_Area),
         Conc_CO2 = as.numeric(Conc_CO2),
         N2O_Area = as.numeric(N2O_Area),
         Conc_N2O = as.numeric(Conc_N2O)
   ) %>% #Ensure Numbers are numeric
  dplyr::select(Sample_ID,
                Type,
                Run_Date,
                CH4_Area,
                Conc_CH4,
                CH4_RTime_Flag,
                CO2_Area,
                Conc_CO2,
                CO2_RTime_Flag,
                N2O_Area,
                Conc_N2O,
                N2O_RTime_Flag,
                FID_Peak_Flag,
                ECD_Peak_Flag) #Select Wanted Columns
  
```

# Assess standard curves CH4
```{r , echo=FALSE}

stds_ch4 <-  raw %>%
  filter(str_detect(Type, "Standard"))

#Correct data class
stds_ch4[, "Conc_CH4"] <- as.numeric(stds_ch4[, "Conc_CH4"])

#Separate Into Low and High

mnmx_stds <- stds_ch4 %>% filter(Conc_CH4 > 0) %>% group_by(Run_Date) %>% mutate(Min = min(Conc_CH4), Max = max(Conc_CH4)) %>% dplyr::select(Run_Date, Min, Max) %>% unique()

low_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 > low_std_crv_cutoff_ch4)

low_std_crv_area_cutoff_CH4 <- low_stds_ch4$CH4_Area[which(low_stds_ch4$Conc_CH4 == max(low_stds_ch4$Conc_CH4))]

low_stds_ch4$Conc_CH4 <- as.numeric(low_stds_ch4$Conc_CH4)


high_stds_ch4 <- stds_ch4

high_stds_ch4$Conc_CH4 <- as.numeric(high_stds_ch4$Conc_CH4)

MidTri_12mL_area <- stds_ch4 %>% filter(Sample_ID %ilike% "12mL" & Sample_ID %ilike% "MidTri") %>% dplyr::select(Sample_ID, CH4_Area, CO2_Area, N2O_Area, Run_Date) %>% rename(MidTri_12mL_CH4_Area = CH4_Area,
                                                   MidTri_12mL_CO2_Area = CO2_Area,                                                                                           MidTri_12mL_N2O_Area = N2O_Area)

#High Curve for CH4
CH4_Curve <- ggplot(high_stds_ch4, aes(Conc_CH4, CH4_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High CH4 Std Curve")
CH4_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_ch4_grouped <- high_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "High",
                                        R2 = summary(lm(CH4_Area ~ Conc_CH4, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ Conc_CH4, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ Conc_CH4, data = .))[1]))

 Slope_CH4_raw <- bind_rows(Slope_CH4_raw_list)
  
 Slope_CH4_High <- Slope_CH4_raw %>% dplyr::select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()


                              
#low Curve for CH4
CH4_Curve <- ggplot(low_stds_ch4, aes(Conc_CH4, CH4_Area, color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low CH4 Std Curve")
CH4_Curve

low_CH4_lm <- lm(low_stds_ch4$CH4_Area ~ low_stds_ch4$Conc_CH4)
low_cf <- coef(low_CH4_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_ch4_grouped <- low_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list_l <- map(split_data_l, ~ .x %>% mutate(Curve = "Low",
                                        R2 = summary(lm(CH4_Area ~ Conc_CH4, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ Conc_CH4, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ Conc_CH4, data = .))[1]))

 Slope_CH4_raw_l <- bind_rows(Slope_CH4_raw_list_l)
  
 Slope_CH4_Low <- Slope_CH4_raw_l %>% dplyr::select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()

 Slope_CH4 <- rbind(Slope_CH4_High, Slope_CH4_Low)
 
 #format dates
 mnmx_stds <- mnmx_stds %>% mutate(Run_Date = paste0(Run_Date))
 
 Slope_CH4 <- Slope_CH4 %>% mutate(Run_Date = paste0(Run_Date)) %>% mutate(Analyte = "CH4")
 
 Slope_CH4_m <- left_join(Slope_CH4, mnmx_stds, by = join_by(Run_Date))

#add new standard curve data to log
log <- read.csv(log_path)
head(log)
log <- rbind(log, Slope_CH4)

#remove possible duplicates from multiple QAQC attempts 

log_CH4 <- log %>% distinct(.keep_all = TRUE)

plot_log <- log_CH4 %>% mutate(Run_Date = parse_date_time(Run_Date, c("mdy","ymd") )) %>%
  filter(Run_Date >= today() - months(12)) %>% filter(Analyte == "CH4")

Slopes_chk <- ggplot(plot_log, aes(x = Run_Date, y = Slope, col = Curve)) +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

Slopes_chk


#check r2

ifelse(Slope_CH4_High$R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_CH4_High$Run_Date), paste0("High_R2 is Good","-", Slope_CH4_High$Run_Date))

ifelse(Slope_CH4_Low$R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_CH4_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_CH4_Low$Run_Date))

High_R2_Flags <- Slope_CH4_High %>% mutate(CH4_High_Flag = ifelse(R2 < r2_cutoff,"High_CH4_R2 is Bad", "")) %>% dplyr::select(CH4_High_Flag, Run_Date)


Low_R2_Flags <- Slope_CH4_Low %>% mutate(CH4_Low_Flag = ifelse(R2 < r2_cutoff,"Low_CH4_R2 is Bad", "")) %>% dplyr::select(CH4_Low_Flag, Run_Date)

Flags_df <- full_join(High_R2_Flags, Low_R2_Flags, by = join_by(Run_Date))

```

# Assess standard curves CO2
```{r , echo=FALSE}

stds_CO2 <-  raw %>%
  filter(str_detect(Type, "Standard"))

#Correct data class
stds_CO2[, "Conc_CO2"] <- as.numeric(stds_CO2[, "Conc_CO2"])

#Separate Into Low and High

mnmx_stds <- stds_CO2 %>% filter(Conc_CO2 > 0) %>% group_by(Run_Date) %>% mutate(Min = min(Conc_CO2), Max = max(Conc_CO2)) %>% dplyr::select(Run_Date, Min, Max) %>% unique()


high_stds_CO2 <- stds_CO2

high_stds_CO2$Conc_CO2 <- as.numeric(high_stds_CO2$Conc_CO2)

#High Curve for CO2
CO2_Curve <- ggplot(high_stds_CO2, aes(Conc_CO2, CO2_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "CO2 Std Curve")
CO2_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_CO2_grouped <- high_stds_CO2 %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_CO2_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_CO2_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "Full",
                                        R2 = summary(lm(CO2_Area ~ Conc_CO2, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CO2_Area ~ Conc_CO2, data = .))[2],
                                        Intercept = coef(lm(CO2_Area ~ Conc_CO2, data = .))[1]))

 Slope_CO2_raw <- bind_rows(Slope_CO2_raw_list)
 
  mnmx_stds <- mnmx_stds %>% mutate(Run_Date = paste0(Run_Date))
  
 Slope_CO2 <- Slope_CO2_raw %>% dplyr::select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct() %>% mutate(Run_Date = paste0(Run_Date)) %>% mutate(Analyte = "CO2")
 
 Slope_CO2_m <- left_join(Slope_CO2, mnmx_stds, by = join_by(Run_Date)) %>% mutate(Run_Date = ymd(Run_Date))

#add new standard curve data to log
log <- log_CH4
head(log)
log <- rbind(log, Slope_CO2)

#remove possible duplicates from multiple QAQC attempts 

log_CO2 <- log %>% distinct(.keep_all = TRUE)

plot_log <- log_CO2 %>% mutate(Run_Date = parse_date_time(Run_Date, c("mdy","ymd"))) %>%
  filter(Run_Date >= today() - months(12)) %>% filter(Analyte == "CO2")

Slopes_chk <- ggplot(plot_log, aes(x = Run_Date, y = Slope, col = Curve)) +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

Slopes_chk

#write.csv(unique_log, log_path, row.names = FALSE)

#check r2

ifelse(Slope_CO2$R2 < r2_cutoff, paste0("CO2_R2 is Bad","-", Slope_CO2$Run_Date), paste0("CO2_R2 is Good","-", Slope_CO2$Run_Date))

R2_Flags_CO2 <- Slope_CO2 %>% mutate(CO2_Flag = ifelse(R2 < r2_cutoff,"CO2_R2 is Bad", "")) %>% dplyr::select(CO2_Flag, Run_Date)

Flags_df <- full_join(R2_Flags_CO2, Flags_df, by = join_by(Run_Date))

```
# Assess standard curves N2O
```{r , echo=FALSE}

stds_N2O <-  raw %>%
  filter(str_detect(Type, "Standard"))

#Correct data class
stds_N2O[, "Conc_N2O"] <- as.numeric(stds_N2O[, "Conc_N2O"])

#Separate Into Low and High

mnmx_stds <- stds_N2O %>% filter(Conc_N2O > 0) %>% group_by(Run_Date) %>% mutate(Min = min(Conc_N2O), Max = max(Conc_N2O)) %>% dplyr::select(Run_Date, Min, Max) %>% unique()

low_stds_N2O <- stds_N2O  %>%
  filter(!Conc_N2O > low_std_crv_cutoff_N2O)

low_std_crv_area_cutoff_N2O <- low_stds_N2O$N2O_Area[which(low_stds_N2O$Conc_N2O == max(low_stds_N2O$Conc_N2O))]

low_stds_N2O$Conc_N2O <- as.numeric(low_stds_N2O$Conc_N2O)


high_stds_N2O <- stds_N2O

high_stds_N2O$Conc_N2O <- as.numeric(high_stds_N2O$Conc_N2O)

#High Curve for N2O
N2O_Curve <- ggplot(high_stds_N2O, aes(Conc_N2O, N2O_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High N2O Std Curve")
N2O_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_N2O_grouped <- high_stds_N2O %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_N2O_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_N2O_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "High",
                                        R2 = summary(lm(N2O_Area ~ Conc_N2O, data = .))$adj.r.squared , 
                                        Slope = coef(lm(N2O_Area ~ Conc_N2O, data = .))[2],
                                        Intercept = coef(lm(N2O_Area ~ Conc_N2O, data = .))[1]))

 Slope_N2O_raw <- bind_rows(Slope_N2O_raw_list)
  
 Slope_N2O_High <- Slope_N2O_raw %>% dplyr::select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()


                              
#low Curve for N2O
N2O_Curve <- ggplot(low_stds_N2O, aes(Conc_N2O, N2O_Area, color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low N2O Std Curve")
N2O_Curve

low_N2O_lm <- lm(low_stds_N2O$N2O_Area ~ low_stds_N2O$Conc_N2O)
low_cf <- coef(low_N2O_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_N2O_grouped <- low_stds_N2O %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_N2O_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_N2O_raw_list_l <- map(split_data_l, ~ .x %>% mutate(Curve = "Low",
                                        R2 = summary(lm(N2O_Area ~ Conc_N2O, data = .))$adj.r.squared , 
                                        Slope = coef(lm(N2O_Area ~ Conc_N2O, data = .))[2],
                                        Intercept = coef(lm(N2O_Area ~ Conc_N2O, data = .))[1]))

 Slope_N2O_raw_l <- bind_rows(Slope_N2O_raw_list_l)
  
 Slope_N2O_Low <- Slope_N2O_raw_l %>% dplyr::select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()
 
 Slope_N2O <- rbind(Slope_N2O_High, Slope_N2O_Low)
 
 #format dates
 mnmx_stds <- mnmx_stds %>% mutate(Run_Date = paste0(Run_Date))
 
 Slope_N2O <- Slope_N2O %>% mutate(Run_Date = paste0(Run_Date)) %>% mutate(Analyte = "N2O")
 
 Slope_N2O_m <- left_join(Slope_N2O, mnmx_stds, by = join_by(Run_Date))

#add new standard curve data to log
log <- log_CO2
head(log)
log <- rbind(log, Slope_N2O)

#remove possible duplicates from multiple QAQC attempts 

unique_log <- log %>% distinct(.keep_all = TRUE)

plot_log <- unique_log %>% mutate(Run_Date = parse_date_time(Run_Date, c("mdy","ymd") )) %>%
  filter(Run_Date >= today() - months(12)) %>% filter(Analyte == "N2O")

Slopes_chk <- ggplot(plot_log, aes(x = Run_Date, y = Slope, col = Curve)) +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

Slopes_chk

#check r2

ifelse(Slope_N2O_High$R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_N2O_High$Run_Date), paste0("High_R2 is Good","-", Slope_N2O_High$Run_Date))

ifelse(Slope_N2O_Low$R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_N2O_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_N2O_Low$Run_Date))

High_R2_Flags_N2O <- Slope_N2O_High %>% mutate(N2O_High_Flag = ifelse(R2 < r2_cutoff,"N2O_R2 is Bad", "")) %>% mutate(Analyte = "N2O") %>% dplyr::select(N2O_High_Flag, Run_Date)


Low_R2_Flags_N2O <- Slope_N2O_Low %>% mutate(N2O_Low_Flag = ifelse(R2 < r2_cutoff,"N2O_R2 is Bad", "")) %>% mutate(Analyte = "N2O") %>% dplyr::select(N2O_Low_Flag, Run_Date)

N2O_Flags_df <- full_join(High_R2_Flags_N2O, Low_R2_Flags_N2O, by = join_by(Run_Date))

Flags_curves <- full_join(N2O_Flags_df, Flags_df, by = join_by(Run_Date))

Flags_df <- Flags_curves %>% mutate(Flag = paste0(CH4_High_Flag, CH4_Low_Flag, CO2_Flag, N2O_High_Flag, N2O_Low_Flag)) %>% dplyr::select(Flag, Run_Date)

```

#Plot Check Areas over Time
```{r ,echo = FALSE}

old_check_log <- read.csv(check_log_path) %>% mutate(Run_Date = ymd(Run_Date))

MT_areas <- stds_ch4 %>% filter(Sample_ID %ilike% "12mL" & Sample_ID %ilike% "MidTri") %>% dplyr::select(CH4_Area, CO2_Area, N2O_Area, Run_Date) %>% rename(MidTri_12mL_CH4_Area = CH4_Area,
                                                   MidTri_12mL_CO2_Area = CO2_Area,                                                                                           MidTri_12mL_N2O_Area = N2O_Area) %>% mutate(Run_Date = ymd(Run_Date))

new_log <- rbind(old_check_log, MT_areas) %>% mutate(Run_Date = ymd(Run_Date))

ch4_area_plot <- ggplot(new_log, aes(x = Run_Date, y = MidTri_12mL_CH4_Area)) +
  labs(title = "CH4 12mL Mid Tri Area") +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

ch4_area_plot

co2_area_plot <- ggplot(new_log, aes(x = Run_Date, y = MidTri_12mL_CO2_Area)) +
  labs(title = "CO2 12mL Mid Tri Area") +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

co2_area_plot

n2o_area_plot <- ggplot(new_log, aes(x = Run_Date, y = MidTri_12mL_N2O_Area)) +
  labs(title = "N2O 12mL Mid Tri Area") +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

n2o_area_plot



```




## Now calculate the CH4 in ppm  
```{r Calculate Concetration CH4, echo=FALSE}
#head(raw)

#pull out Samples
manual_ch4 <- Manual_Areas %>%
  filter(Analyte == "CH4") %>%
  select(Sample_ID = Sample, Manual_CH4_Area = Area) %>%
  mutate(Manual_CH4_Area = as.numeric(Manual_CH4_Area))

Samples_1 <- raw %>%  
  filter(str_detect(Type, "Sample")) %>%
  filter(!is.na(CH4_Area)) %>%
  filter(!is.na(CO2_Area)) %>%
  filter(!is.na(N2O_Area)) %>%
  left_join(manual_ch4, by = "Sample_ID") %>%
  mutate(CH4_Area = coalesce(Manual_CH4_Area, CH4_Area)) %>%
  mutate(Curve = ifelse(CH4_Area >= low_std_crv_area_cutoff_CH4, "High", "Low")) %>%
  select(-Manual_CH4_Area)

#head(Samples)


Slope_CH4_m <-Slope_CH4_m %>% mutate(Run_Date = ymd(Run_Date))

Samples_2 <- left_join(Samples_1, Slope_CH4_m, by = c("Run_Date", "Curve"))

#Calculate CH4 concentrations in ppm
Samples_CH4 <- Samples_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area - Intercept) / Slope)

##########make flags for any dilutions needed 


Samples_CH4 <- Samples_CH4 %>% mutate(CH4_Flag = ifelse(CH4_Conc_ppm > Max, "ADL", ifelse(CH4_Conc_ppm < Min, "BDL", "Within Range"))) 

Samples_CH4 <- Samples_CH4 %>% dplyr::select(-R2, 
                                            -Slope,
                                            -Intercept,
                                            -Analyte,
                                            -Min,
                                            -Max,
                                            -Curve)

#head(Samples)

```



## Now calculate the CO2 in ppm  
```{r Calculate Concetration CO2, echo=FALSE}
#head(raw)
manual_CO2 <- Manual_Areas %>%
  filter(Analyte == "CO2") %>%
  select(Sample_ID = Sample, Manual_CO2_Area = Area) %>%
  mutate(Manual_CO2_Area = as.numeric(Manual_CO2_Area))

Samples_2 <- full_join(Samples_CH4, Slope_CO2_m, by = join_by(Run_Date)) %>% 
  left_join(manual_CO2, by = "Sample_ID") %>%
  mutate(CO2_Area = coalesce(Manual_CO2_Area, CO2_Area))%>%
  select(-Manual_CO2_Area)

#Calculate CO2 concentrations in ppm
Samples <- Samples_2 %>% 
  mutate(CO2_Conc_ppm = (CO2_Area - Intercept) / Slope)

##########make flags for any dilutions needed 


Samples <- Samples %>% mutate(CO2_Flag = ifelse(CO2_Conc_ppm > Max, "ADL", ifelse(CO2_Conc_ppm < Min, "BDL", "Within Range")))
  
Samples_CO2 <- Samples %>% dplyr::select(-R2, 
                                            -Slope,
                                            -Intercept,
                                            -Analyte,
                                            -Min,
                                            -Max,
                                            -Curve)

#head(Samples)

```

## Now calculate the N2O in ppm  
```{r Calculate Concetration N2O, echo=FALSE}
#head(raw)


manual_N2O <- Manual_Areas %>%
  filter(Analyte == "N2O") %>%
  select(Sample_ID = Sample, Manual_N2O_Area = Area) %>%
  mutate(Manual_N2O_Area = as.numeric(Manual_N2O_Area))

Slope_N2O_m <-Slope_N2O_m %>% mutate(Run_Date = ymd(Run_Date))

Samples_2 <- Samples_CO2 %>% mutate(Curve = ifelse(N2O_Area >= low_std_crv_area_cutoff_N2O, "High", "Low")) %>% left_join(Slope_N2O_m, by = c("Run_Date", "Curve")) %>% 
  left_join(manual_N2O, by = "Sample_ID") %>%
  mutate(N2O_Area = coalesce(Manual_N2O_Area, N2O_Area))%>%
  select(-Manual_N2O_Area)

#Calculate N2O concentrations in ppm
Samples_N2O <- Samples_2 %>% 
  mutate(N2O_Conc_ppm = (N2O_Area - Intercept) / Slope)

##########make flags for any dilutions needed 

Samples_N2O <- Samples_N2O %>% mutate(N2O_Flag = ifelse(N2O_Conc_ppm > Max, "ADL", ifelse(N2O_Conc_ppm < Min, "BDL", "Within Range"))) 

Samples <- Samples_N2O %>% dplyr::select(-R2, 
                                             -Slope,
                                             -Intercept,
                                             -Analyte,
                                             -Min,
                                             -Max,
                                             -Curve)

#head(Samples)

```

# Aanlayze the CH4 Check Standards 
```{r Asses CH4 Check Standards, echo=FALSE}

chks_1 <-  raw %>%
  filter(str_detect(Type, "ChkStd")) %>%
  mutate(Curve = "High") %>%
  filter(!is.na(CH4_Area))

#head(chks)

chks_2 <- merge(chks_1, Slope_CH4)


chks <- chks_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area - Intercept) / Slope)

#calculate percent difference of check standards
chks$Ch4_diff <- ((chks$CH4_Conc_ppm - level_ch4) / ((chks$CH4_Conc_ppm + level_ch4) /
                                                       2)) * 100
chks$CH4_diff_flag <-  ifelse(abs(chks$Ch4_diff) < 10, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CH4_chks_percent <- (sum(chks$CH4_diff_flag == "YES") / nrow(chks)) * 100

#report out if flags indicate need for rerun

if (CH4_chks_percent >= chks_flag) {
  message(">60% of CH4 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CH4 Check Standards are within range of expected concentration - REASSESS")
}

#now plot the Ch4 concentrations and the CO2 concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

ch4_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CH4_Conc_ppm, fill =
                                      CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: CH4") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_ch4,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


ch4_chk

#Add Check Standard Flags

flag_chk <- ifelse(CH4_chks_percent >= chks_flag,
                "",
               "; CH4 Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk))

```

# Aanlayze the CO2 Check Standards 
```{r Asses CO2 Check Standards, echo=FALSE}

chks_1 <-  raw %>%
  filter(str_detect(Type, "ChkStd")) %>%
  mutate(Curve = "Full") %>%
  filter(!is.na(CO2_Area))

#head(chks)

chks_2 <- merge(chks_1, Slope_CO2)


chks <- chks_2 %>% 
  mutate(CO2_Conc_ppm = (CO2_Area - Intercept) / Slope)

#calculate percent difference of check standards
chks$CO2_diff <- ((chks$CO2_Conc_ppm - level_co2) / ((chks$CO2_Conc_ppm + level_co2) /
                                                       2)) * 100
chks$CO2_diff_flag <-  ifelse(abs(chks$CO2_diff) < 10, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CO2_chks_percent <- (sum(chks$CO2_diff_flag == "YES") / nrow(chks)) * 100

#report out if flags indicate need for rerun

if (CO2_chks_percent >= chks_flag) {
  message(">60% of CO2 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CO2 Check Standards are within range of expected concentration - REASSESS")
}

#now plot the CO2 concentrations and the CO2 concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

CO2_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CO2_Conc_ppm, fill =
                                      CO2_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: CO2") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_co2,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


CO2_chk

#Add Check Standard Flags

flag_chk <- ifelse(CO2_chks_percent >= chks_flag,
                "",
               "; CO2 Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk))

```


# Aanlayze the N2O Check Standards 
```{r Asses N2O Check Standards, echo=FALSE}

chks_1 <-  raw %>%
  filter(str_detect(Type, "ChkStd")) %>%
  mutate(Curve = "High") %>%
  filter(!is.na(N2O_Area))

#head(chks)

chks_2 <- merge(chks_1, Slope_N2O)


chks <- chks_2 %>% 
  mutate(N2O_Conc_ppm = (N2O_Area - Intercept) / Slope)

#calculate percent difference of check standards
chks$N2O_diff <- ((chks$N2O_Conc_ppm - level_n2o) / ((chks$N2O_Conc_ppm + level_n2o) /
                                                       2)) * 100
chks$N2O_diff_flag <-  ifelse(abs(chks$N2O_diff) < 10, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
N2O_chks_percent <- (sum(chks$N2O_diff_flag == "YES") / nrow(chks)) * 100

#report out if flags indicate need for rerun

if (N2O_chks_percent >= chks_flag) {
  message(">60% of N2O Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of N2O Check Standards are within range of expected concentration - REASSESS")
}

#now plot the N2O concentrations and the N2O concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

N2O_chk <-  ggplot(data = chks, aes(x = Run_Date, y = N2O_Conc_ppm, fill =
                                      N2O_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: N2O") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_n2o,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


N2O_chk

#Add Check Standard Flags

flag_chk <- ifelse(N2O_chks_percent >= chks_flag,
                "",
               "; N2O Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk))

```

# Analyze the CH4 Blanks 
```{r Analyze CH4 Blanks, echo=FALSE}
cat("Assess CH4 Blanks")

CH4_blks <-  raw %>%
  filter(str_detect(Type, "Blank")) %>%
  filter(!is.na(CH4_Area))


#Pull out samples from df_all to calc quantile
 samples_blanks <- raw %>%
  filter(str_detect(Type, c("Sample")))

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

blk_flag_CH4 <- quantile(samples_blanks$CH4_Area, prob = c(.25), na.rm = TRUE)   #this gives you the lower 25% quartile of the data
CH4_blks$CH4_diff_flag <-  ifelse(CH4_blks$CH4_Area <= blk_flag_CH4, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CH4_blks_percent <- (sum(CH4_blks$CH4_diff_flag == "YES") / nrow(CH4_blks)) *
  100

#report out if flags indicate need for rerun
if (CH4_blks_percent >= blks_flag) {
  message(">60% of CH4 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CH4 Blanks are lower 25% quartile of samples - REASSESS")
}

#plotting the blanks compared to the lower 25% of conc (what the flag is)
CH4_blks_plot <-  ggplot(data = CH4_blks, aes(x = Run_Date, y = CH4_Area, fill =
                                                CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CH4 Area", title = "CH4 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CH4),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(CH4_blks_plot)


#find average of run blanks for flagging samples later
# Compute averages
blk_avg_CH4 <- mean(CH4_blks$CH4_Area, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(Test = c("CH4"),
                         Blank_Mean_Area = c(blk_avg_CH4))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks <-  ifelse(CH4_blks_percent >= blks_flag,
                 "",
                 '; Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks))




```

# Analyze the CO2 Blanks 
```{r Analyze CO2 Blanks, echo=FALSE}
cat("Assess CO2 Blanks")

CO2_blks <-  raw %>%
  filter(str_detect(Type, "Blank"))%>%
  filter(!is.na(CO2_Area))


#Pull out samples from df_all to calc quantile
 samples_blanks <- raw %>%
  filter(str_detect(Type, c("Sample")))

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

blk_flag_CO2 <- quantile(samples_blanks$CO2_Area, prob = c(.25), na.rm = TRUE)   #this gives you the lower 25% quartile of the data
CO2_blks$CO2_diff_flag <-  ifelse(CO2_blks$CO2_Area <= blk_flag_CO2, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CO2_blks_percent <- (sum(CO2_blks$CO2_diff_flag == "YES") / nrow(CO2_blks)) *
  100

#report out if flags indicate need for rerun
if (CO2_blks_percent >= blks_flag) {
  message(">60% of CO2 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CO2 Blanks are lower 25% quartile of samples - REASSESS")
}

#plotting the blanks compared to the lower 25% of conc (what the flag is)
CO2_blks_plot <-  ggplot(data = CO2_blks, aes(x = Run_Date, y = CO2_Area, fill =
                                                CO2_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CO2 Area", title = "CO2 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CO2),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(CO2_blks_plot)


#find average of run blanks for flagging samples later
# Compute averages
blk_avg_CO2 <- mean(CO2_blks$CO2_Area, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(Test = c("CO2"),
                         Blank_Mean_Area = c(blk_avg_CO2))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks <-  ifelse(CO2_blks_percent >= blks_flag,
                 "",
                 '; Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks))




```

# Analyze the N2O Blanks 
```{r Analyze N2O Blanks, echo=FALSE}
cat("Assess N2O Blanks")

N2O_blks <-  raw %>%
  filter(str_detect(Type, "Blank"))%>%
  filter(!is.na(N2O_Area))


#Pull out samples from df_all to calc quantile
 samples_blanks <- raw %>%
  filter(str_detect(Type, c("Sample")))

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

blk_flag_N2O <- quantile(samples_blanks$N2O_Area, prob = c(.25), na.rm = TRUE)   #this gives you the lower 25% quartile of the data
N2O_blks$N2O_diff_flag <-  ifelse(N2O_blks$N2O_Area <= blk_flag_N2O, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
N2O_blks_percent <- (sum(N2O_blks$N2O_diff_flag == "YES") / nrow(N2O_blks)) *
  100

#report out if flags indicate need for rerun
if (N2O_blks_percent >= blks_flag) {
  message(">60% of N2O Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of N2O Blanks are lower 25% quartile of samples - REASSESS")
}

#plotting the blanks compared to the lower 25% of conc (what the flag is)
N2O_blks_plot <-  ggplot(data = N2O_blks, aes(x = Run_Date, y = N2O_Area, fill =
                                                N2O_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "N2O Area", title = "N2O Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_N2O),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(N2O_blks_plot)


#find average of run blanks for flagging samples later
# Compute averages
blk_avg_N2O <- mean(N2O_blks$N2O_Area, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(Test = c("N2O"),
                         Blank_Mean_Area = c(blk_avg_N2O))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks <-  ifelse(N2O_blks_percent >= blks_flag,
                 "",
                 '; Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks))




```

# Check Range - CH4
```{r, echo=FALSE}

#check results
#head(Samples)

#quick first look at the samples
ch4_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm, fill =
                                             CH4_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "bottom")


ch4_samples



```

# Check Range - CO2
```{r, echo=FALSE}

#check results
#head(Samples)

#quick first look at the samples
CO2_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CO2_Conc_ppm, fill =
                                             CO2_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "CO2 (ppm)", title = "CO2: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "bottom")

CO2_samples



```

# Check Range - N2O
```{r, echo=FALSE}

#check results
#head(Samples)

#quick first look at the samples
N2O_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = N2O_Conc_ppm, fill =
                                             N2O_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "N2O (ppm)", title = "N2O: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "bottom")


N2O_samples




```


# Calculate gas in water - CH4
```{r, echo=FALSE}

# Convert ppm CH4 to total uM CH4 in the sample headspace
# use Ideal Gas Law:
#   hdsp umol CH4 = ppm CH4*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

#Calculate CH4 uM from ppm

#Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

#old equation
#Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * ((1 * volume / 1000) / (R * (TempC + CTK)))) / (volume / 1000)

Samples <- Samples %>% mutate(CH4_Conc_uM = ifelse(GIW == "YES" | Sample_ID %in% GIW_Samples, (CH4_Conc_ppm * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w, NA))


#quick first look at the samples
ch4_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "CH4 (uM)", title = "CH4 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


ch4_samples1

```

# Calculate gas in water - CO2
```{r, echo=FALSE}

# Convert ppm CO2 to total uM CO2 in the sample headspace
# use Ideal Gas Law:
#   hdsp umol CO2 = ppm CO2*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

#Calculate CO2 uM from ppm

#Samples$CO2_Conc_uM <- (Samples$CO2_Conc_ppm * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

#old equation
#Samples$CO2_Conc_uM <- (Samples$CO2_Conc_ppm_dilcorr * ((1 * volume / 1000) / (R * (TempC + CTK)))) / (volume / 1000)

Samples <- Samples %>% mutate(CO2_Conc_uM = ifelse(GIW == "YES" | Sample_ID %in% GIW_Samples, (CO2_Conc_ppm * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w, NA))


#quick first look at the samples
CO2_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = CO2_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "CO2 (uM)", title = "CO2 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


CO2_samples1

```

# Calculate gas in water - N2O
```{r, echo=FALSE}

# Convert ppm N2O to total uM N2O in the sample headspace
# use Ideal Gas Law:
#   hdsp umol N2O = ppm N2O*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

#Calculate N2O uM from ppm

#Samples$N2O_Conc_uM <- (Samples$N2O_Conc_ppm * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

#old equation
#Samples$N2O_Conc_uM <- (Samples$N2O_Conc_ppm_dilcorr * ((1 * volume / 1000) / (R * (TempC + CTK)))) / (volume / 1000)

Samples <- Samples %>% mutate(N2O_Conc_uM = ifelse(GIW == "YES" | Sample_ID %in% GIW_Samples, (N2O_Conc_ppm * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w, NA))


#quick first look at the samples
N2O_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = N2O_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "N2O (uM)", title = "N2O Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


N2O_samples1

```

# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata
all_present <- all(Clean_Metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(Clean_Metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}



#merge metadata with sample run data
merged_data <- Clean_Metadata %>% rename(Sample_ID = Sample_Name) %>% left_join(Samples, by = join_by(Sample_ID))

#might want to include a flag or something to see that all the samples are included?

#add flags

merged_2 <- merge(merged_data, Flags_df)
merged_data <-merged_2 %>% mutate(Run_Flags = Flag)



df_all_clean <- merged_data

```

## Organize Data  
```{r, echo=FALSE}
#check results
  head(Samples)
 
  #pull out what we need
  df_all_clean_cols <- df_all_clean %>%
    select(
      Sample1,
      Sample_Year,
      Sample_Month,
      Sample_Day,
      Sample_Time,
      Time_Zone,
      G_W,
      Sample_Plot,
      Sample_Grid,
      Sample_Depth,
      Field_Notes,
      Lab_Notes,
      CH4_Conc_ppm,
      CH4_Conc_uM,
      CH4_Flag,
      CO2_Conc_ppm,
      CO2_Conc_uM,
      CO2_Flag,
      N2O_Conc_ppm,
      N2O_Conc_uM,
      N2O_Flag,
      Run_Flags,
      Run_Date
    )
 
  final_data_labeled <- df_all_clean_cols %>%
    mutate(
      Run_notes = run_notes, #new column with notes about the run
      Analysis_rundate = Run_Date)
 
  #rename columns
  org_labeled_fin_data <- final_data_labeled %>%
    rename(
      CH4_Conc_Flag = CH4_Flag,
      QAQC_Flags = Run_Flags,
      CO2_Conc_Flag = CO2_Flag,
      N2O_Conc_Flag = N2O_Flag,
      
    )
 
  #Order Columns for Final Data Frame
  All_Clean_Data <- org_labeled_fin_data %>%
    select(
     Sample1,
     Sample_Year,
     Sample_Month,
     Sample_Day,
     Sample_Time,
     Time_Zone,
     G_W,
     Sample_Plot,
     Sample_Grid,
     Sample_Depth,
     Field_Notes,
     Lab_Notes,
     CH4_Conc_ppm,
     CH4_Conc_uM,
     CH4_Conc_Flag,
     CO2_Conc_ppm,
     CO2_Conc_uM,
     CO2_Conc_Flag,
     N2O_Conc_ppm,
     N2O_Conc_uM,
     N2O_Conc_Flag,
     QAQC_Flags,
     Analysis_rundate,
     Run_notes
    )
  head(All_Clean_Data)

```



## Visualize Data CH4
```{r Visualize Data CH4, fig.height=10, echo=FALSE}
cat("Visualize Data CH4")

### Ammonia 
CH4_forplot <- All_Clean_Data

#group the data for plotting
CH4_forplot <- CH4_forplot %>%
  group_by(Sample_Depth) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup() %>%
  mutate(Depth1 = ifelse(Sample_Depth == "ATM", "Atmospheric", paste0(Sample_Depth, " cm"))) %>% mutate(Depth = factor(Depth1, levels = c("Atmospheric", "10 cm", "20 cm", "35 cm", "50 cm")))

#create plot
viz_CH4_plot <- ggplot(data = CH4_forplot, aes(
  x = as.factor(Sample_Grid),
  y = CH4_Conc_ppm,
  fill = as.factor(Sample_Plot)
)) +
  geom_col(stat = "identity", 
           position = position_dodge2(padding = 0.1, 
                                      width = 0.5, 
                                      preserve =  "single" )) +
  facet_grid(Depth ~ Sample_Plot, 
             scales = "free_y") +
  scale_fill_manual(
    breaks = c(
       "Control", "Saltwater", "Freshwater"),
    values = c("Control" = "springgreen2", "Freshwater" = "cyan2",
                                      "Saltwater" = "violetred2"),
    name = "Plot:"
  ) +
  theme_classic() +
  labs(
    x = " ",
    y = "CH4 Conc (ppm)",
    title = "TEMPEST: Soil Gas Well CH4"
  ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE, limits = rev) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8),
      legend.key = element_rect("black"),
    legend.position = "none"
    ) + 
  coord_flip()
print(viz_CH4_plot)

```

## Visualize Data CO2
```{r Visualize Data CO2, fig.height=10, echo=FALSE}
cat("Visualize Data CO2")

### Ammonia 
CO2_forplot <- All_Clean_Data

#group the data for plotting
CO2_forplot <- CO2_forplot %>%
  group_by(Sample_Depth) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup() %>%
  mutate(Depth1 = ifelse(Sample_Depth == "ATM", "Atmospheric", paste0(Sample_Depth, " cm"))) %>% mutate(Depth = factor(Depth1, levels = c("Atmospheric", "10 cm", "20 cm", "35 cm", "50 cm")))

#create plot
viz_CO2_plot <- ggplot(data = CO2_forplot, aes(
  x = as.factor(Sample_Grid),
  y = CO2_Conc_ppm,
  fill = as.factor(Sample_Plot)
)) +
  geom_col(stat = "identity", 
           position = position_dodge2(padding = 0.1, 
                                      width = 0.5, 
                                      preserve =  "single" )) +
  facet_grid(Depth ~ Sample_Plot, 
             scales = "free_y") +
  scale_fill_manual(
    breaks = c(
       "Control", "Saltwater", "Freshwater"),
    values = c("Control" = "springgreen2", "Freshwater" = "cyan2",
                                      "Saltwater" = "violetred2"),
    name = "Plot:"
  ) +
  theme_classic() +
  labs(
    x = " ",
    y = "CO2 Conc (ppm)",
    title = "TEMPEST: Soil Gas Well CO2"
  ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE, limits = rev) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8),
      legend.key = element_rect("black"),
    legend.position = "none"
    ) + 
  coord_flip()
print(viz_CO2_plot)

```
\newpage

## Visualize Data N2O
```{r Visualize Data N2O, fig.height=10, echo=FALSE}
cat("Visualize Data N2O")

### Ammonia 
N2O_forplot <- All_Clean_Data

#group the data for plotting
N2O_forplot <- N2O_forplot %>%
  group_by(Sample_Depth) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup() %>%
  mutate(Depth1 = ifelse(Sample_Depth == "ATM", "Atmospheric", paste0(Sample_Depth, " cm"))) %>% mutate(Depth = factor(Depth1, levels = c("Atmospheric", "10 cm", "20 cm", "35 cm", "50 cm")))

#create plot
viz_N2O_plot <- ggplot(data = N2O_forplot, aes(
  x = as.factor(Sample_Grid),
  y = N2O_Conc_ppm,
  fill = as.factor(Sample_Plot)
)) +
  geom_col(stat = "identity", 
           position = position_dodge2(padding = 0.1, 
                                      width = 0.5, 
                                      preserve =  "single" )) +
  facet_grid(Depth ~ Sample_Plot, 
             scales = "free_y") +
  scale_fill_manual(
    breaks = c(
       "Control", "Saltwater", "Freshwater"),
    values = c("Control" = "springgreen2", "Freshwater" = "cyan2",
                                      "Saltwater" = "violetred2"),
    name = "Plot:"
  ) +
  theme_classic() +
  labs(
    x = " ",
    y = "N2O Conc (ppm)",
    title = "TEMPEST: Soil Gas Well N2O"
  ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE, limits = rev) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8),
      legend.key = element_rect("black"),
    legend.position = "none"
    ) + 
  coord_flip()
print(viz_N2O_plot)

```
\newpage

#Visualize Data CH4 by Replicate
```{r, echo=FALSE}
#ensure points are in correct order

CH4_forplot_1 <- CH4_forplot %>% filter(Sample_Depth != "ATM") %>% mutate(Depth_cm = as.numeric(Sample_Depth))

#make plot
viz_CH4_plot_2 <- ggplot(data = CH4_forplot_1, aes(x = CH4_Conc_ppm, y = as.numeric(Depth_cm), color = Sample_Grid)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  facet_grid(Sample_Plot ~ ., scales = "fixed") +
  labs(
    x = "CH4 Conc (ppm)",
    y = "Depth (cm)",
    title = "TEMPEST: Soil Gas Well CH4 by Replicate",
    color = "Sample Grid:"
  ) +
  scale_color_manual(
    values = c(
      "C3" = "#FFD500",
      "C6" = "#00D0FF",
      "F4" = "#FF0075",
      "H3" = "#80d280",
      "H6" = "#ff6040"
    )) +
  scale_y_reverse(breaks = c(10, 20, 35, 50), limits = c(55, 5)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_CH4_plot_2
```

#Visualize Data CO2 by Replicate
```{r, echo=FALSE}
#ensure points are in correct order

CO2_forplot_1 <- CO2_forplot %>% filter(Sample_Depth != "ATM") %>% mutate(Depth_cm = as.numeric(Sample_Depth))

#make plot
viz_CO2_plot_2 <- ggplot(data = CO2_forplot_1, aes(x = CO2_Conc_ppm, y = as.numeric(Depth_cm), color = Sample_Grid)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  facet_grid(Sample_Plot ~ ., scales = "fixed") +
  labs(
    x = "CO2 Conc (ppm)",
    y = "Depth (cm)",
    title = "TEMPEST: Soil Gas Well CO2 by Replicate",
    color = "Sample Grid:"
  ) +
  scale_color_manual(
    values = c(
      "C3" = "#FFD500",
      "C6" = "#00D0FF",
      "F4" = "#FF0075",
      "H3" = "#80d280",
      "H6" = "#ff6040"
    )) +
  scale_y_reverse(breaks = c(10, 20, 35, 50), limits = c(55, 5)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_CO2_plot_2
```

#Visualize Data N2O by Replicate
```{r, echo=FALSE}
#ensure points are in correct order

N2O_forplot_1 <- N2O_forplot %>% filter(Sample_Depth != "ATM") %>% mutate(Depth_cm = as.numeric(Sample_Depth))

#make plot
viz_N2O_plot_2 <- ggplot(data = N2O_forplot_1, aes(x = N2O_Conc_ppm, y = as.numeric(Depth_cm), color = Sample_Grid)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  facet_grid(Sample_Plot ~ ., scales = "fixed") +
  labs(
    x = "N2O Conc (ppm)",
    y = "Depth (cm)",
    title = "TEMPEST: Soil Gas Well N2O by Replicate",
    color = "Sample Grid:"
  ) +
  scale_color_manual(
    values = c(
      "C3" = "#FFD500",
      "C6" = "#00D0FF",
      "F4" = "#FF0075",
      "H3" = "#80d280",
      "H6" = "#ff6040"
    )) +
  scale_y_reverse(breaks = c(10, 20, 35, 50), limits = c(55, 5)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_N2O_plot_2
```



##Export Data
```{r, echo=FALSE}
#Write Data
write.csv(All_Clean_Data, final_path, row.names = FALSE)

#Write Logs
write.csv(unique_log, log_path, row.names = FALSE)

write.csv(new_log, check_log_path, row.names = FALSE)
#end
```

#end